---
title: "HIC: Gut Microbiota - Heat Tolerance Analysis (manuscript)"
author: "Alex Gould"
date: "`r date()`"
output:
  pdf_document:
    toc: yes
    toc_depth: 4
    number_sections: no
    latex_engine: xelatex
  rmarkdown::pdf_document:
    extra_dependencies: float
    toc: yes
html_document:
  toc: yes
toc_float: yes
---

# Section 0: Setup

## Global options

```{r global options, include = TRUE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      error = FALSE,
                      eval = TRUE,
                      dev = c('pdf', 'png'),
                      width = 60,
                      fig.align = 'center', 
                      fig.height = 10, 
                      fig.width = 10, 
                      fig.dpi = 300,
                      fig.pos = "!H",
                      dpi = 300,
                      out.extra = "",
                      message = FALSE, 
                      cache = FALSE,
                      pdf.options(encoding = "ISOLatin9.enc"),
                      fig.path='figs/rmd/') 
```

## Parameters

To run via knitting directly in RStudio, set 'eval = TRUE' in the below chunk, and add your parameter values below:

``` {r SET_parameters, eval = TRUE}
analysis_dir     = '...'
qiime_dir        = '...'
data_dir         = '...'
pw_message       = "...."
experiment_type  = 'illumina'               
load_type        = 'input'
experiment_code  = '...'
process_num      = '1'
indir            = '...'
min_otus         = 20000
nclass           = 15 
group_cols       = c('Tolerant' = '#005900', 'Intolerant' = '#ff2c15') 
custom_col16     = c(
  "#771155",
  "#AA4488",
  "#CC99BB",
  "#114477",
  "#4477AA",
  "#117777",
  "#77CCCC",
  "#117744",
  "#88CCAA",
  "#777711",
  "#AAAA44",
  "#DDDD77",
  "#774411",
  "#AA7744",
  "#DDAA77",
  "#771122"
)
test_type        = "non-parametric"
control          = "Tolerant"
otu_or_asv       = "ASV"
read_type        = "filter"
rep_shapes       = rep(c(15, 16, 17, 18, 3, 7, 8, 9, 10, 11, 12, 13), 5)
```

## Environment setup
First check that the input parameters are suitable:

``` {r check_parameters}
if (!file.exists(analysis_dir)) stop("Error: Output directory ", analysis_dir, " not found. Please ensure that the output directory exists before running.\n")
if (!experiment_type %in% c("illumina", "nanopore")) stop ("Error: experiment_type must be one of 'illumina' or 'nanopore'.\n")
if (!load_type %in% c("pipeline", "input")) stop ("Error: load_type must be one of 'pipeline' or 'input'.\n")
if(load_type == "pipeline" && (is.na(experiment_code) || is.na(process_num))) stop("Error: Parameters 'experiment_code' and 'process_num' are required if load_type 'pipeline' used.\n")
if(load_type == "input" && (is.na(indir))) stop("Error: Parameters 'experiment_code' and 'process_num' are required if load_type 'pipeline' used.\n")
if(load_type == "input" && (!file.exists(indir))) stop("Error: Input directory ", indir, " not found. Please ensure that the input directory containing the required input files is correct before running.\n")
if (nclass < 5 || nclass > 20) stop("Error: nclass must be between 5 and 20.\n")
if (min_otus < 5000) stop("Error: min_otus must be greater than 5000.\n")
if (!test_type %in% c("parametric", "non-parametric")) stop ("Error: test_type must be one of 'parametric' or 'non-parametric'.\n")
if (!read_type %in% c("raw", "trim", "filter")) stop ("Error: read_type must be one of 'raw', 'trim' or 'filter'.\n")
###!!!! Check group_cols and control
if(!file.exists(paste0(analysis_dir, "/tabs"))) dir.create(paste0(analysis_dir, "/tabs"))
```

Then set up the R environment:

```{r setup}
## Load the libraries
library("DBI")
library("RMySQL")
library("phyloseq")
library("tidyverse")
library("xlsx")
library("rstudioapi")
library("qiime2R")
library("jmv")
library("ggpubr")
library("ggtext")
library("vegan")
library("plyr")
library("reshape2")
library("DESeq2")
library("broom") 
library("scales") 
library("corrplot") 
library("openxlsx")
library("Hmisc")
library("knitr")
library("ggvenn")
library("rlang")
library("ALDEx2")
library("ANCOMBC")

## Set seed for reproducibility
set.seed(0)

## Set font for figures
font = element_text(family = "serif")
```

## stat_compare_means() fix

Fix for `stat_compare_means()` function to allow Times New Roman font. Thank you https://stackoverflow.com/questions/77346644/stat-compare-means-unable-to-change-font-family-when-using-comparisons-argu 

```{r stat_compare_means_fix}
fixed_call <- quote(ggsignif::geom_signif(comparisons = comparisons, 
                                          y_position = label.y, 
                                          test = method, test.args = method.args, 
                                          step_increase = step.increase, 
                                          size = bracket.size, textsize = 6, color = color, 
                                          map_signif_level = map_signif_level, 
                                          tip_length = tip.length, data = data,
                                          vjust = vjust, ...))

scm <- as.list(ggpubr::stat_compare_means)
scm[[26]][[2]][[3]][[13]] <- fixed_call
stat_compare_means <- function(...) {
  ggp <- getNamespace('ggpubr')
  .method_info <- ggp$.method_info
  .add_item <- ggp$.add_item
  .is_p.signif_in_mapping <- ggp$.is_p.signif_in_mapping
  .is_empty <- ggp$.is_empty
  do.call(as.function(scm), list(...), quote = FALSE)
}
```

## Load data

### Sample annotation

``` {r load_sample_ann}
## load sample annotation here... 
```

### Set sample annotation

``` {r SET_sample_ann}
## set sample annotation here...
```

### Phyloseq

``` {r load_phyloseq_data}
## load phyloseq here...
```

# Section 1: Quality Control - all data

## Rarefaction curve

``` {r Section1_QC_alldata_rarefaction_curve, cache = FALSE}
## Calculate a rarefaction curve (code taken from https://github.com/joey711/phyloseq/issues/143)
calculate_rarefaction_curves <- function(psdata, measures, depths) {

  estimate_rarified_richness <- function(psdata, measures, depth) {
    if(max(sample_sums(psdata)) < depth) return()
    psdata <- prune_samples(sample_sums(psdata) >= depth, psdata)

    rarified_psdata <- rarefy_even_depth(psdata, depth, verbose = FALSE)

    alpha_diversity <- estimate_richness(rarified_psdata, measures = measures)

    ## as.matrix forces the use of melt.array, which includes the Sample names (rownames)
    molten_alpha_diversity <- melt(as.matrix(alpha_diversity), varnames = c('Sample', 'Measure'), value.name = 'Alpha_diversity')

    molten_alpha_diversity
  }

  names(depths) <- depths ## this enables automatic addition of the Depth to the output by ldply
  rarefaction_curve_data <- ldply(depths, estimate_rarified_richness, psdata = psdata, measures = measures, .id = 'Depth', .progress = ifelse(interactive(), 'text', 'none'))

  ## convert Depth from factor to numeric
  rarefaction_curve_data$Depth <- as.numeric(levels(rarefaction_curve_data$Depth))[rarefaction_curve_data$Depth]

  rarefaction_curve_data
}

## Calculate the rarefaction curve
set.seed(42)
rarefaction_curve_data <- calculate_rarefaction_curves(phyloseq_dat,
                                                       c('Observed', "Chao1", 'Shannon', "Simpson"),
                                                       rep(c(1, 10, 100, 1000, 1:100 * 1000), each = 10))

## Summarize the alpha diversity
rarefaction_curve_data_summary <- ddply(rarefaction_curve_data, c('Depth', 'Sample', 'Measure'),
                                        summarise,
                                        Alpha_diversity_mean = mean(Alpha_diversity, na.rm = TRUE),
                                        Alpha_diversity_sd = sd(Alpha_diversity, na.rm = TRUE))
rarefaction_curve_data_summary [["Sample"]] <- gsub("\\.", " ", rarefaction_curve_data_summary [["Sample"]])
rarefaction_curve_data_summary_verbose <- merge(rarefaction_curve_data_summary, data.frame(sample_data(phyloseq_dat)), by.x = 'Sample', by.y = 'row.names')
rarefaction_curve_data_summary_verbose[["Depth"]] <- rarefaction_curve_data_summary_verbose[["Depth"]]/1000
```

### Plot

``` {r Section1_QC_alldata_rarefaction_curve_plot, cache = FALSE, fig.height = 10, fig.width = 15}
## Plot the rarefaction curve
rarefaction_curve <- ggplot(data = subset(rarefaction_curve_data_summary_verbose, Measure %in% c('Observed', "Chao1", 'Shannon', "Simpson")),
                            mapping = aes(
                              x = Depth,
                              y = Alpha_diversity_mean,
                              ymin = Alpha_diversity_mean - Alpha_diversity_sd,
                              ymax = Alpha_diversity_mean + Alpha_diversity_sd,
                              colour = Group,
                              shape  = Rep,
                              #group  = Rep,
                              alpha = 0.8)) +
  geom_line() +
  geom_pointrange(size = 1) +
  xlab("Read Depth (1,000s)") +
  ylab("Alpha Diversity") +
  scale_fill_manual(values = group_cols) +
  scale_color_manual(values = group_cols) +
  scale_shape_manual(values = rep_shapes) +
  facet_grid(Measure ~ ., scales="free_y") +
  theme_bw() +
  theme(
    text = font,
    axis.title = element_text(size = 24),
    axis.text = element_text(size = 18),
    legend.text = element_text(size = 18),
    legend.title = element_text(size=18),
    strip.text.x = element_text(size=24, face = "bold"),
    strip.text.y = element_text(size=24, face = "bold"))

rarefaction_curve

## reset seed to 0
set.seed(0)
```

## Taxa count per sample - All data

The following plot shows the total number of observed taxa (OTUs or ASVs) per sample, with the minimum number of taxa specified above (min_otus) highlighted with a dashed line. Samples lower than this will be removed below in Section 2. 

``` {r Section1_QC_alldata_OTU_count, fig.height = 12, fig.width = 10}
data.frame(sample_data(phyloseq_dat)) %>%
  dplyr::mutate(count = sample_sums(phyloseq_dat)) %>%
  ggplot(aes(x = UIN, y = count/1000, fill = Group)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  xlab("") + ylab(paste0("Total Number ", otu_or_asv, "s (x1000)")) +
  facet_wrap(.~Group, scales = "free_x") +
  geom_hline(yintercept = min_otus/1000, linetype = 2, linewidth = 1.2) +
  scale_fill_manual(values = group_cols) +
  scale_color_manual(values = group_cols) +
  scale_shape_manual(values = rep_shapes) +
  theme_bw() +
  theme(
    text = font,
    axis.title   = element_text(size = 18),
    axis.text.y  = element_text(size = 12),
    axis.text.x  = element_text(size = 14, angle = 315, face = "bold", hjust = 0),
    legend.text  = element_text(size = 14),
    legend.title = element_text(size = 20, face = "bold"),
    strip.text.x = element_text(size=20, face = "bold"),
    strip.text.y = element_text(size=16, face = "bold")
  )
```

## Sample filtering and rarefaction

In this section, the data will be filtered to remove all samples with fewer than `r min_otus` OTUs in the data set. The remaining data will be rarefied to reduce all samples to the same depth, which will be the lowest depth from the remaining samples following filtering.

``` {r Section1_QC_filterdata_filter_and_rarefy_samples}
## Filter to remove samples with fewer than min_otus
phyloseq_dat_filter <- prune_samples(sample_sums(phyloseq_dat) >= min_otus, phyloseq_dat)

## Create a non rarefied phyloseq object with filtered samples for DESeq2 analysis
phyloseq_dat_deseq <- prune_samples(sample_sums(phyloseq_dat) >= min_otus, phyloseq_dat)

## Rarefy the data to the minimum otu depth
rarefy_otu_depth <- min(sample_sums(phyloseq_dat_filter))
cat("Rarefication of data to", rarefy_otu_depth, "OTUs:\n")
phyloseq_dat_filter <- rarefy_even_depth(phyloseq_dat_filter, rarefy_otu_depth, verbose = FALSE)
```

### Summary N for Analysis

Below, a summary of the n taken forward for analysis is provided: 

``` {r Section1_n_summary}
table(sample_ann_summary$Group)
```

## Taxa count per sample - Rarefied and filtered data

The following plot shows the total number of observed taxa (OTUs or ASVs) per sample after rarefaction and filtering of samples to remove low coverage samples, with the minimum number of taxa specified above (min_otus) highlighted with a dashed line. Samples lower than this will be removed below in Section 2. All samples should have the same depth. If this is not the case, then please debug the code.

``` {r Section1_QC_filterdata_OTU_count, fig.height = 12, fig.width = 10}
data.frame(sample_data(phyloseq_dat_filter)) %>%
  dplyr::mutate(count = sample_sums(phyloseq_dat_filter)) %>%
  ggplot(aes(x = UIN, y = count/1000, fill = Group)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  xlab("") + ylab(paste0("Total Number ", otu_or_asv, "s (x1000)")) +
  facet_wrap(.~Group, scales = "free_x") +
  geom_hline(yintercept = min_otus/1000, linetype = 2, linewidth = 1.2) +
  scale_fill_manual(values = group_cols) +
  scale_color_manual(values = group_cols) +
  scale_shape_manual(values = rep_shapes) +
  theme_bw() +
  theme(
    text = font,
    axis.title   = element_text(size = 18),
    axis.text.y  = element_text(size = 12),
    axis.text.x  = element_text(size = 14, angle = 315, face = "bold", hjust = 0),
    legend.text  = element_text(size = 14),
    legend.title = element_text(size = 20, face = "bold"),
    strip.text.x = element_text(size=20, face = "bold"),
    strip.text.y = element_text(size=16, face = "bold")
  )
```

# Section 2: Microbial Diversity

## Alpha diversity - Rarefied and filtered data

```{r Section2_QC_filterdata_alpha_diversity_summary}
## Create alpha diversity data frame
adiv <- data.frame(
  "UIN" = phyloseq::sample_data(phyloseq_dat_filter)$UIN,
  "Rep" = phyloseq::sample_data(phyloseq_dat_filter)$Rep,
  "Group" = phyloseq::sample_data(phyloseq_dat_filter)$Group,
  "Observed" = phyloseq::estimate_richness(phyloseq_dat_filter, measures = "Observed"),
  "Chao1" = phyloseq::estimate_richness(phyloseq_dat_filter, measures = "Chao1"),
  "Shannon" = phyloseq::estimate_richness(phyloseq_dat_filter, measures = "Shannon"),
  "Simpson" = phyloseq::estimate_richness(phyloseq_dat_filter, measures = "Simpson"))

adiv <- adiv %>%
  ## Remove Chao1 SE column
  dplyr::select(-Chao1.se.chao1) %>%
  ## Rename Chao1 column
  dplyr::rename(Chao1 = Chao1.Chao1)

## write as csv
write.table(
  adiv,
  paste0(analysis_dir, "/tabs/alpha_diversity_table.csv"),
  row.names = FALSE,
  col.names = TRUE,
  quote = FALSE,
  sep = ","
)
``` 

### Normality

```{r Section2_QC_filterdata_alpha_diversity_normality}
jmv::descriptives(
    formula = Observed + Chao1 + Shannon + Simpson ~ Group,
    data = adiv,
    iqr = TRUE,
    sw = TRUE)
```

### t-test

```{r Section2_QC_filterdata_alpha_diversity_comparison_stat}
jmv::ttestIS(
    formula = Observed + Chao1 + Shannon + Simpson ~ Group,
    data = adiv,
    vars = vars(Observed, Chao1, Shannon, Simpson),
    eqv = TRUE,
    effectSize = TRUE)
```

### Plot

```{r Section2_QC_filterdata_alpha_diversity_comparison_plot, fig.height = 8, fig.width = 12}
## Plot alpha diversity
adiv_outcome_plot <- adiv %>%
  tidyr::gather(key = metric, value = value, c("Observed", "Chao1", "Shannon", "Simpson")) %>%
  dplyr::mutate(metric = factor(metric, levels = c("Observed", "Chao1", "Shannon", "Simpson"))) %>%
  ggplot(aes(x = Group, y = value)) +
  # geom_boxplot(outlier.color = NA,
  #              size = 1) +
  geom_jitter(aes(color = Group),
              size = 2,
              alpha = 0.8,
              width = 0.25) +
  stat_summary(fun = mean, 
               geom = "point", 
               color = "black", 
               size = 3, 
               show.legend = FALSE) +
  stat_summary(fun.data = mean_sd, 
               geom = "errorbar", 
               color = "black", 
               width = 0.3, 
               linewidth = 0.8, 
               show.legend = FALSE) +
  labs(x = "") +
  ylab("Alpha Diversity") +
  facet_wrap(~ metric, scales = "free", ncol = 4) +
  stat_compare_means(comparisons = list(c("Tolerant", "Intolerant")),
                     method = "t.test", 
                     label = "p.signif",
                     paired = FALSE,
                     family = "serif",
                     bracket.size = 0.6,
                     vjust = -0.2) +
  scale_fill_manual(values = group_cols) +
  scale_color_manual(values = group_cols) +
  theme_classic() +
  theme(
    text = font,
    axis.title   = element_text(size = 20),
    axis.text.y  = element_text(size = 16),
    axis.text.x  = element_text(size = 12),
    strip.text = element_text(size=26, face = "bold"),
    legend.position = "none") +
  expand_limits(y=0)

## view
adiv_outcome_plot
``` 

## Beta diversity - Rarefied and filtered data

The following principal coordinate analysis (PCoA) plot shows the beta diversity following rarefaction and filtering of samples, based on the Bray Curtis distance between samples. 

``` {r Section2_QC_filterdata_beta_diversity, fig.height = 7, fig.width = 9.5}
## beta diversity - PCoA
bdiv_outcome_plot <- plot_ordination(phyloseq_dat_filter, 
                                     ordinate(phyloseq_dat_filter, 
                                              method = "PCoA", 
                                              distance = "bray", 
                                              na.rm = TRUE), 
                                     type = "samples", 
                                     color = "Group", 
                                     #shape = "Rep"
                                     ) + 
  stat_ellipse(aes(group = Group), 
               linetype = 2,
               linewidth = 1) +
  geom_point(size = 5,
             alpha = 0.8) +
  ggtitle("") + 
  scale_fill_manual(values = group_cols) +
  scale_color_manual(values = group_cols) +
  #scale_shape_manual(values = rep_shapes) +
  theme_classic() +
  theme(
    text = font,
    axis.title = element_text(size = 20),
    axis.text = element_text(size = 16),
    legend.text = element_text(size = 18),
    legend.title = element_text(size=22))

## view
bdiv_outcome_plot
```

## PERMANOVA

This section will use permutational MANOVA (PERMANOVA) to look for significant effects of the Group variable on the abundance. Dissimilarity is again based on the Bray Curtis difference. 

``` {r Section2_PERMANOVA_analysis_Patient_vs_Control}
## Set seed for reproducibility 
set.seed(0)

## Convert the counts across all samples into a community data matrix
plot_dat <- psmelt(phyloseq_dat_filter)
community_mat <- plot_dat %>%
  dplyr::select(Sample, OTU, Abundance) %>%
  spread(OTU, Abundance)
row.names(community_mat) <- community_mat[["Sample"]]
community_mat <- as.matrix(community_mat[, -1]) ## Remove the Sample column

## Transform the data using square root to minimize influence of most abundant groups
community_mat <- sqrt(community_mat)

## Create a dissimilarity matrix
OTU_dist <- vegdist(community_mat, method = "bray")

## Get the matching sample data
ind_dat <- plot_dat %>%
  dplyr::select(Sample, Group) %>%
  distinct()
ind_dat <- ind_dat[match(rownames(community_mat), ind_dat[["Sample"]]), ]

## Run perMANOVA
perm <- adonis2(OTU_dist ~ Group, data = ind_dat, permutations = 9999, method = "bray")
perm
```

# Section 3: Microbial Abundance

## Normalise

Here we normalise the taxa counts to represent the percentage of all taxas.

``` {r Section3_normalise}
phyloseq_dat_filter_norm <- transform_sample_counts(phyloseq_dat_filter, function (x) 100* ((x+1) / sum(x+1)))
phyloseq_dat_filter_norm_melt <- psmelt(phyloseq_dat_filter_norm)

## melted object of rarefied data
phyloseq_dat_filter_melt <- psmelt(phyloseq_dat_filter)

## melted phyloseq object on non-rarefied data for DESeq2
phyloseq_dat_norm_melt <- psmelt(phyloseq_dat_deseq)
```

## Summerise 

``` {r Section3_summarise_taxa}
## rarefied data: print the number and each unique taxas. Each number total will contain one NA. 
length(unique(phyloseq_dat_filter_melt$Phylum))
unique(phyloseq_dat_filter_melt$Phylum)
length(unique(phyloseq_dat_filter_melt$Class))
unique(phyloseq_dat_filter_melt$Class)
length(unique(phyloseq_dat_filter_melt$Order))
unique(phyloseq_dat_filter_melt$Order)
length(unique(phyloseq_dat_filter_melt$Family))
unique(phyloseq_dat_filter_melt$Family)
length(unique(phyloseq_dat_filter_melt$Genus))
unique(phyloseq_dat_filter_melt$Genus)
```

```{r Section3_summarise_taxa_new}
## Function to calculate distinct counts of taxa for each UIN within Group
summarise_taxa <- function(data) {
  data %>%
    ## Filter by a count >= 1
    dplyr::filter(Abundance >= 1) %>%
    ## Group by UIN and Group
    dplyr::group_by(UIN, Group) %>%
    # Summarise unique counts of taxa at each taxonomic level
    dplyr::summarise(
      Phylum = n_distinct(Phylum, na.rm = TRUE),
      Class = n_distinct(Class, na.rm = TRUE),
      Order = n_distinct(Order, na.rm = TRUE),
      Family = n_distinct(Family, na.rm = TRUE),
      Genus = n_distinct(Genus, na.rm = TRUE),
      Species = n_distinct(Species, na.rm = TRUE),
      ASV = n_distinct(OTU, na.rm = TRUE),
      ## drop groupings
      .groups = 'drop'
    )
}

## Return results for unique taxa
summary_taxa_result <- summarise_taxa(phyloseq_dat_filter_melt)

## View the results
head(summary_taxa_result)

## descriptives with S-W test
jmv::descriptives(
    formula = . ~ Group,
    data = summary_taxa_result,
    iqr = TRUE,
    sw = TRUE)

## summarise average & variance of taxa for each taxa level, by group
summary_by_group <- summary_taxa_result %>%
  dplyr::group_by(Group) %>%
  dplyr::summarise(
    ## phyla: median (IQR)
    median_phyla = median(Phylum, na.rm = FALSE),
    iqr_phyla = IQR(Phylum, na.rm = FALSE),
    ## classes: median (IQR)
    median_classes = median(Class, na.rm = FALSE),
    iqr_classes = IQR(Class, na.rm = FALSE),
    ## orders: mean SD
    mean_orders = mean(Order, na.rm = FALSE),
    sd_orders = sd(Order, na.rm = FALSE),
    ## families: mean SD
    mean_families = mean(Family, na.rm = FALSE),
    sd_families = sd(Family, na.rm = FALSE),
    ## genera: mean SD
    mean_genera = mean(Genus, na.rm = FALSE),
    sd_genera = sd(Genus, na.rm = FALSE),
    ## ASV: mean SD
    mean_ASV = mean(ASV, na.rm = FALSE),
    sd_ASV = sd(ASV, na.rm = FALSE)
  ) %>%
  ## round numbers to 0 decimal places
  dplyr::mutate_if(is.numeric, round, digits = 0)

summary_by_group

## Write this into a file
write.table(
  summary_by_group,
  paste0(analysis_dir, "/tabs/taxa_summary_table_by_group.csv"),
  row.names = FALSE,
  col.names = TRUE,
  sep = ",",
  quote = FALSE
)
```

## Barplot function: Normal text

The following function will plot a barplot showing the normalised abundance across all taxonomic levels with normal text for taxa. 

``` {r Section3_plot_barplot_normal_text}
## Function for non-italicised taxa names in the legend
plot_barplot <-
  function (taxa, plot_dat = phyloseq_dat_filter_norm_melt) {
    # taxa = "Genus"
    # plot_dat = phyloseq_dat_filter_norm_melt
    
    ## Get the total abundance across all samples for each OTU
    tmp <- plot_dat %>%
      dplyr::group_by(OTU) %>%
      dplyr::summarise(sum = sum(Abundance))
    tsums <- tmp[["sum"]]
    names(tsums) <- tmp[["OTU"]]
    tsums <- tsums[order(names(tsums))]
    
    ## Get the OTU taxa
    tmp <- subset(plot_dat, Sample == plot_dat[1, "Sample"])
    taxlist <- tmp[, taxa]
    names(taxlist) <- tmp[, "OTU"]
    
    ## Check the data match
    if (length(tsums) != length(taxlist) ||
        !all(names(tsums) %in% names(taxlist)) ||
        !all(names(taxlist) %in% names(tsums))) {
      stop("Error: tsums and taxlist do not match.\n")
    }
    taxlist <- taxlist[names(tsums)]
    
    ## Keep only the top taxa and add all other taxa into an "other" class
    taxasum <-
      sort(tapply(tsums, taxlist, sum, na.rm = TRUE), decreasing = TRUE)
    taxasum <- taxasum[!names(taxasum) == ""]
    plot_dat[["TaxName"]] <- as.character(plot_dat[[taxa]])
    plot_dat[is.na(plot_dat[["TaxName"]]) |
               plot_dat[["TaxName"]] == "", "TaxName"] <- "Other"
    plot_dat[["TaxName"]] <-
      ifelse(plot_dat[["TaxName"]] %in% names(taxasum)[1:nclass],
             as.character(plot_dat[[taxa]]),
             "Other")
    plot_dat[["TaxName"]] <-
      factor(plot_dat[["TaxName"]], levels = c(names(taxasum)[1:nclass], "Other"))
    
    ## Combine counts at the taxa level
    plot_dat <- plot_dat %>%
      dplyr::group_by(Sample, Group, UIN, TaxName) %>%
      dplyr::summarise(Abundance = sum(Abundance)) %>%
      ## remove "[]", "-", "_" from taxa names
      dplyr::mutate(TaxName = gsub("\\[|\\]|-|_", " ", TaxName)) %>%
      ## remove leading and trialing spaces!
      dplyr::mutate(TaxName = trimws(TaxName))
    
    ## Extract unique and edited TaxName values
    unique_taxa <- unique(plot_dat$TaxName)
    
    ## reset factor levels again
    plot_dat <- plot_dat %>%
      dplyr::mutate(TaxName = factor(TaxName, levels = unique_taxa))
    
    ## Write this into a file
    write.table(
      plot_dat,
      paste0(
        analysis_dir,
        "/tabs/individual_abundance_table_",
        taxa,
        ".csv"
      ),
      row.names = FALSE,
      col.names = TRUE,
      sep = ",",
      quote = FALSE
    )
    
    ## Create a colour pallet for the barplot
    set.seed(8)
    cols <-
      sample(custom_col16, length(unique(plot_dat[["TaxName"]])))
    names(cols) <- unique(plot_dat[["TaxName"]])
    cols["Other"] <- "#D3D3D3" ## set other as grey
    
    ## Create the plot with the custom legend formatting
    p <- plot_dat %>%
      ggplot(aes(x = UIN, y = Abundance, fill = TaxName)) +
      geom_bar(stat = "identity", width = 1) +
      labs(x = "",
           y = "Relative Abundance") +
      scale_y_continuous(labels = scales::label_percent(scale = 1),
                         expand = expansion(add = c(0, 1))) +
      scale_fill_manual(values = cols) +
      facet_wrap(. ~ Group, scales = "free_x") +
      theme_classic() +
      theme(
        text = font,
        axis.title = element_text(size = 18),
        axis.text.y = element_text(size = 16),
        axis.text.x = element_blank(),
        legend.text = element_markdown(size = 14),
        ## Format legend text as markdown so names are italicised
        legend.title = element_blank(),
        strip.text.x = element_text(size = 16, face = "bold"),
        strip.text.y = element_text(size = 16, face = "bold")
      )
    
    p
    
    ## Return the figure and the data set
    return(list(p, plot_dat))
}
```

## Barplot function: Italicised Text

The following function will plot a barplot showing the normalised abundance across all taxonomic levels with italicised text for taxa. 

```{r Section3_plot_barplot_italised_text}
## Function for italicised taxa names in the legend
plot_barplot2 <- function (taxa,
                           plot_dat = phyloseq_dat_filter_norm_melt,
                           specific_name = NULL) {
  # taxa = "Genus"
  # plot_dat = phyloseq_dat_NFW_filter_norm_melt
  # specific_name = "negative_control"
  
  ## Construct the csv filename based on whether a specific name is provided
  if (is.null(specific_name)) {
    filename <- paste0("individual_abundance_table_", taxa, ".csv")
  } else {
    filename <-
      paste0("individual_abundance_table_",
             specific_name,
             "_",
             taxa,
             ".csv")
  }
  
  ## Include the analysis directory in the file path
  filepath <- file.path(analysis_dir, "tabs", filename)
  
  ## Get the total abundance across all samples for each OTU
  tmp <- plot_dat %>%
    dplyr::group_by(OTU) %>%
    dplyr::summarise(sum = sum(Abundance))
  tsums <- tmp[["sum"]]
  names(tsums) <- tmp[["OTU"]]
  tsums <- tsums[order(names(tsums))]
  
  ## Get the OTU taxa
  tmp <- subset(plot_dat, Sample == plot_dat[1, "Sample"])
  taxlist <- tmp[, taxa]
  names(taxlist) <- tmp[, "OTU"]
  
  ## Check the data match
  if (length(tsums) != length(taxlist) ||
      !all(names(tsums) %in% names(taxlist)) ||
      !all(names(taxlist) %in% names(tsums))) {
    stop("Error: tsums and taxlist do not match.\n")
  }
  taxlist <- taxlist[names(tsums)]
  
  ## Keep only the top taxa and add all other taxa into an "other" class
  taxasum <-
    sort(tapply(tsums, taxlist, sum, na.rm = TRUE), decreasing = TRUE)
  taxasum <- taxasum[!names(taxasum) == ""]
  plot_dat[["TaxName"]] <- as.character(plot_dat[[taxa]])
  plot_dat[is.na(plot_dat[["TaxName"]]) |
             plot_dat[["TaxName"]] == "", "TaxName"] <- "Other"
  plot_dat[["TaxName"]] <-
    ifelse(plot_dat[["TaxName"]] %in% names(taxasum)[1:nclass],
           as.character(plot_dat[[taxa]]),
           "Other")
  plot_dat[["TaxName"]] <-
    factor(plot_dat[["TaxName"]], levels = c(names(taxasum)[1:nclass], "Other"))
  
  ## Combine counts at the taxa level
  plot_dat <- plot_dat %>%
    dplyr::group_by(Sample, Group, UIN, TaxName) %>%
    dplyr::summarise(Abundance = sum(Abundance)) %>%
    dplyr::mutate(## remove "[]", "-", "_" from taxa names
      TaxName = gsub("\\[|\\]|-|_", " ", TaxName),
      ## remove leading and trialing spaces!
      TaxName = trimws(TaxName)) %>%
    ## remove instances of doubles spaces
    dplyr::mutate(TaxName = gsub("  ", " ", TaxName)) %>%
    ## Italicising taxa names
    dplyr::mutate(FormattedTaxName = ifelse(
      ## keep uncultured as normal text
      grepl("^uncultured", TaxName),
      gsub("^(uncultured) (.+)$", "\\1 *\\2*", TaxName),
      ## Keep the stated names (e.g. "Other") as normal text (i.e. not italicised) and wrap taxa names with "*"
      ifelse(
        TaxName %in% c(
          "Other",
          "metagenome",
          "gut metagenome",
          "human gut",
          "unidentified"
        ),
        TaxName,
        paste("*", TaxName, "*", sep = "")
      )
    )) %>%
    ## insert line break in long taxa names
    dplyr::mutate(
      FormattedTaxName = ifelse(
        FormattedTaxName == "*Eubacterium hallii group*",
        "*Eubacterium<br>hallii group*",
        FormattedTaxName
      )
    )
  
  ## Extract unique and edited TaxName values
  unique_taxa <- unique(plot_dat$TaxName)
  unique_taxa2 <- unique(plot_dat$FormattedTaxName)
  
  ## reset factor levels again
  plot_dat <- plot_dat %>%
    dplyr::mutate(
      TaxName = factor(TaxName, levels = unique_taxa),
      FormattedTaxName = factor(FormattedTaxName, levels = unique_taxa2)
    )
  
  ## Write this into a file
  write.table(
    plot_dat,
    file = filepath,
    row.names = FALSE,
    col.names = TRUE,
    sep = ",",
    quote = FALSE
  )
  
  ## Create a colour pallet for the barplot
  set.seed(8)
  cols <-
    sample(custom_col16, length(unique(plot_dat[["FormattedTaxName"]])))
  names(cols) <- unique(plot_dat[["FormattedTaxName"]])
  cols["Other"] <- "#D3D3D3" ## set other as grey
  
  ## Create the plot with the custom legend formatting
  p <- plot_dat %>%
    ggplot(aes(x = UIN, y = Abundance, fill = FormattedTaxName)) +
    geom_bar(stat = "identity", width = 1) +
    labs(x = "",
         y = "Relative Abundance") +
    scale_y_continuous(labels = scales::label_percent(scale = 1),
                       expand = expansion(add = c(0, 1))) +
    scale_fill_manual(values = cols) +
    facet_wrap(. ~ Group, scales = "free_x") +
    theme_classic() +
    theme(
      text = font,
      axis.title = element_text(size = 18),
      axis.text.y = element_text(size = 16),
      axis.text.x = element_blank(),
      legend.text = element_markdown(size = 14),
      ## Format legend text as markdown so names are italicised
      legend.title = element_blank(),
      strip.text.x = element_text(size = 16, face = "bold"),
      strip.text.y = element_text(size = 16, face = "bold")
    )
  
  p
  
  ## Return the figure and the data set
  return(list(p, plot_dat))
}
```

### Phylum
``` {r Section3_barplot_top_phylum_normalised_abundance_stacked, fig.height = 8, fig.width = 8}
phylum_dat <- plot_barplot("Phylum")
phylum_p <- phylum_dat[[1]]
phylum_p
```

### Class
``` {r Section3_barplot_top_class_normalised_abundance_stacked, fig.height = 8, fig.width = 8}
class_dat <- plot_barplot("Class")
class_dat[[1]]
```

### Order
``` {r Section3_barplot_top_order_normalised_abundance_stacked, fig.height = 8, fig.width = 8}
order_dat <- plot_barplot("Order")
order_dat[[1]]
```

### Family
``` {r Section3_barplot_top_family_normalised_abundance_stacked, fig.height = 8, fig.width = 8}
family_dat <- plot_barplot2("Family")
family_dat[[1]]
```

### Genus
``` {r Section3_barplot_top_genus_normalised_abundance_stacked, fig.height = 8, fig.width = 8}
genus_dat <- plot_barplot2("Genus")
genus_p <- genus_dat[[1]]
genus_p
```

## Firmicutes/Bacteroidota ratio

```{r Section3_calculate_FBratio}
fbratio_outcome <- phylum_dat[[2]] %>%
  dplyr::group_by(UIN) %>%
  tidyr::spread(TaxName, Abundance) %>%
  dplyr::mutate(FBratio = Firmicutes/Bacteroidota) %>%
  dplyr::select(UIN, Group, Firmicutes, Bacteroidota, FBratio) 

## Write as csv
write.table(
  fbratio_outcome,
  paste0(analysis_dir, "/tabs/FBratio_table.csv"),
  row.names = FALSE,
  col.names = TRUE,
  sep = ",",
  quote = FALSE
)

head(fbratio_outcome)
```

### Normality

```{r FBratio_normality}
jmv::descriptives(
    formula = FBratio ~ Group,
    data = fbratio_outcome,
    iqr = TRUE,
    sw = TRUE)
```

### t-test

```{r FBratio_comparison_stat}
jmv::ttestIS(
    formula = FBratio ~ Group,
    data = fbratio_outcome,
    vars = FBratio,
    students = FALSE,
    mann = TRUE,
    effectSize = TRUE)
```

### Plot

``` {r Section3_FBratio_plot, fig.height = 6, fig.width = 6}
## plot figure
set.seed(123)
fbratio_outcome_plot <- fbratio_outcome %>% 
  ggplot(aes(x = Group, y = FBratio)) +
  geom_boxplot(outlier.color = NA,
               size = 1) +
  geom_jitter(aes(color = Group),
              size = 2,
              alpha = 0.8) +
  labs(x = "", 
       y = "F:B Ratio") +
  stat_compare_means(comparisons = list(c("Tolerant", "Intolerant")),
                     method = "wilcox.test",
                     label = "p.signif",
                     paired = FALSE,
                     family = "serif",
                     bracket.size = 0.6,
                     vjust = -0.2) +
  scale_y_continuous(breaks = seq(0, 20, 2), 
                     limits = c(0, 20)) +
  scale_color_manual(values = group_cols) +
  theme_classic() +
  theme(
    text = font,
    axis.title   = element_text(size = 18),
    axis.text.y  = element_text(size = 16),
    axis.text.x  = element_text(size = 18),
    legend.position = "none"
    ) 

## view
fbratio_outcome_plot
```

## Abundance Comparison All Taxa Function

Below, we will compare all taxa, at each taxonomic level between groups. 

``` {r Section3_abundance_comparison_function}
## Function for non-italicised taxa names
compare_abundance_comparison <- function (taxa,
                                          plot_dat = phyloseq_dat_filter_norm_melt,
                                          p_type = "individual") {
  # taxa = "Phylum"
  # plot_dat = phyloseq_dat_filter_norm_melt
  # p_type = "individual"
  
  ## Get the total abundance across all samples for each OTU for the given taxa
  tmp <- plot_dat %>%
    dplyr::group_by(OTU) %>%
    dplyr::summarise(sum = sum(Abundance))
  tsums <- tmp[["sum"]]
  names(tsums) <- tmp[["OTU"]]
  tsums <- tsums[order(names(tsums))]
  
  ## Get the OTU taxa
  tmp <- subset(plot_dat, Sample == plot_dat[1, "Sample"])
  taxlist <- tmp[, taxa]
  names(taxlist) <- tmp[, "OTU"]
  
  ## Check the data match
  if (length(tsums) != length(taxlist) ||
      !all(names(tsums) %in% names(taxlist)) ||
      !all(names(taxlist) %in% names(tsums))) {
    stop("Error: tsums and taxlist do not match.\n")
  }
  taxlist <- taxlist[names(tsums)]
  
  taxasum <-
    sort(tapply(tsums, taxlist, sum, na.rm = TRUE), decreasing = TRUE)
  taxasum <- taxasum[!names(taxasum) == ""]
  plot_dat[["TaxName"]] <- as.character(plot_dat[[taxa]])
  plot_dat[["TaxName"]] <- factor(plot_dat[[taxa]])
  
  ## Combine counts at the taxa level
  plot_dat <- plot_dat %>%
    dplyr::group_by(UIN, Group, TaxName) %>%
    dplyr::summarise(Abundance = sum(Abundance), .groups = "drop") %>%
    dplyr::filter(TaxName != "Other")
  
  ## create separate wide data frame for all taxa of each individual UIN
  UIN_taxa_wide <- plot_dat
  UIN_taxa_wide <- UIN_taxa_wide %>%
    tidyr::pivot_wider(names_from = TaxName,
                       values_from = Abundance)
  ## save as .csv file
  write.table(
    UIN_taxa_wide,
    paste0(
      analysis_dir,
      "/tabs/individual_relative_abundance_",
      taxa,
      ".csv"
    ),
    sep = ",",
    row.names = FALSE,
    col.names = TRUE,
    quote = FALSE
  )
  
  ## Combine at the Group level
  plot_dat_SE <- plot_dat %>%
    dplyr::group_by(Group, TaxName) %>%
    dplyr::summarise(
      average = ifelse(
        test_type == "parametric",
        mean(Abundance),
        median(Abundance)
      ),
      variability = ifelse(
        test_type == "parametric",
        sd(Abundance) / sqrt(length(Abundance)),
        IQR(Abundance)
      )
    )
  
  pval <-
    data.frame(TaxName       = levels(plot_dat_SE[["TaxName"]]),
               pval          = NA)
  for (t in pval[["TaxName"]]) {
    if (test_type == "parametric") {
      pval[pval[["TaxName"]] == t, "pval"] <-
        anova(lm(Abundance ~ Group, data = subset(plot_dat, TaxName == t)))[1, "Pr(>F)"]
    } else if (test_type == "non-parametric") {
      pval[pval[["TaxName"]] == t, "pval"] <-
        kruskal.test(Abundance ~ Group, data = subset(plot_dat, TaxName == t))[["p.value"]]
    }
  }
  
  ## Calculate individual p-values vs control
  for (t in pval[["TaxName"]]) {
    for (g in unique(plot_dat[["Group"]])) {
      if (g == "Control") {
        pval[pval[["TaxName"]] == t, paste0("pval.", g)] <- 1
      } else if (test_type == "parametric") {
        pval[pval[["TaxName"]] == t, paste0("pval.", g)] <-
          t.test(
            subset(plot_dat, TaxName == t &
                     Group == g)[["Abundance"]],
            subset(plot_dat, TaxName == t &
                     Group == control)[["Abundance"]]
          )[["p.value"]]
      } else if (test_type == "non-parametric") {
        pval[pval[["TaxName"]] == t, paste0("pval.", g)] <-
          wilcox.test(
            subset(plot_dat, TaxName == t &
                     Group == g)[["Abundance"]],
            subset(plot_dat, TaxName == t &
                     Group == control)[["Abundance"]]
          )[["p.value"]]
      }
    }
  }
  
  ## Add to data frame
  plot_dat_SE[["pval"]] <- NA
  for (t in pval[["TaxName"]]) {
    for (g in unique(plot_dat[["Group"]])) {
      if (p_type == "combine") {
        if (g == control) {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <-
            pval[pval[["TaxName"]] == t, "pval"]
        } else {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <- NA
        }
      } else if (p_type == "individual") {
        if (g == control) {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <- NA
        } else {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <-
            pval[pval[["TaxName"]] == t, paste0("pval.", g)]
        }
      }
    }
  }
  
  ## Correct for multiple testing
  which_nonna <- which(!is.na(plot_dat_SE[["pval"]]))
  plot_dat_SE[["padj"]] <- NA
  plot_dat_SE[which_nonna, "padj"] <-
    p.adjust(as.numeric(plot_dat_SE[which_nonna,][["pval"]]),
             method = "fdr",
             n = nrow(plot_dat_SE[which_nonna,]))
  
  ## Convert to star format
  plot_dat_SE[["padj_star"]] <-
    case_when(
      plot_dat_SE[["padj"]] <= 0.001 ~ "***",
      plot_dat_SE[["padj"]] <= 0.01  ~ "**",
      plot_dat_SE[["padj"]] <= 0.05  ~ "*",
      TRUE ~ ""
    )
  
  plot_dat_SE
  
  ## Save the values
  write.table(
    plot_dat_SE,
    paste0(
      analysis_dir,
      "/tabs/abundance_comparison_table_",
      taxa,
      ".csv"
    ),
    sep = ",",
    row.names = FALSE,
    col.names = TRUE,
    quote = FALSE
  )
  
  ## format table
  plot_dat_format <- plot_dat_SE %>%
    tidyr::pivot_wider(
      names_from = Group,
      values_from = c(average, variability, pval, padj, padj_star)
    ) %>%
    dplyr::mutate(
      average_Tolerant = sprintf("%.1f", average_Tolerant),
      variability_Tolerant = sprintf("%.1f", variability_Tolerant),
      average_Intolerant = sprintf("%.1f", average_Intolerant),
      variability_Intolerant = sprintf("%.1f", variability_Intolerant),
      Tolerant = paste0(average_Tolerant, " (", variability_Tolerant, ")"),
      Intolerant = paste0(average_Intolerant, " (", variability_Intolerant, ")")
    ) %>%
    dplyr::rename(pval = pval_Intolerant,
                  padj = padj_Intolerant,
                  Taxa = TaxName) %>%
    dplyr::mutate(pval = sprintf("%.3f", as.numeric(pval)),
                  padj = sprintf("%.3f", as.numeric(padj))) %>%
    dplyr::select(Taxa, Tolerant, Intolerant, pval, padj)
  
  ## save formatted table
  write.table(
    plot_dat_format,
    paste0(
      analysis_dir,
      "/tabs/abundance_comparison_table_formatted_",
      taxa,
      ".csv"
    ),
    sep = ",",
    row.names = FALSE,
    col.names = TRUE,
    quote = FALSE
  )
  
  return(plot_dat_format)
}
```

### Phylum

```{r Section3_abundance_comparison_Phylum}
phylum_comparison <- compare_abundance_comparison("Phylum")
phylum_comparison

## filter by pval
phylum_comparison %>%
  dplyr::filter(pval < 0.05)

## filter by padj
phylum_comparison %>%
  dplyr::filter(padj < 0.05)
```

### Class

```{r Section3_abundance_comparison_Class}
class_comparison <- compare_abundance_comparison("Class")
class_comparison

## filter by pval
class_comparison %>%
  dplyr::filter(pval < 0.05)

## filter by padj
class_comparison %>%
  dplyr::filter(padj < 0.05)
```

### Order

```{r Section3_abundance_comparison_Order}
order_comparison <- compare_abundance_comparison("Order")
order_comparison

## filter by pval
order_comparison %>%
  dplyr::filter(pval < 0.05)

## filter by padj
order_comparison %>%
  dplyr::filter(padj < 0.05)
```

### Family

```{r Section3_abundance_comparison_Family}
family_comparison <- compare_abundance_comparison("Family")
family_comparison

## filter by pval
family_comparison %>%
  dplyr::filter(pval < 0.05)

## filter by padj
family_comparison %>%
  dplyr::filter(padj < 0.05)
```

### Genus

```{r Section3_abundance_comparison_Genus}
genus_comparison <- compare_abundance_comparison("Genus")
genus_comparison

## filter by pval
genus_comparison %>%
  dplyr::filter(pval < 0.05)

## filter by padj
genus_comparison %>%
  dplyr::filter(padj < 0.05)
```

## Plot Abundance Comparison Function:

### Abundance Comparison Function: Normal Text

The following function will plot the difference between the groups at different taxonomic levels with normal text for taxa:

``` {r Section3_plot_abundance_comparison_normal_text}
## Function for non-italicised taxa names
plot_abundance_comparison <- function (taxa,
                                       plot_dat = phyloseq_dat_filter_norm_melt,
                                       p_type = "individual",
                                       nclass = 15) {
  # taxa = "Genus"
  # plot_dat = phyloseq_dat_filter_norm_melt
  # p_type = "individual"
  
  ## Get the total abundance across all samples for each OTU for the given taxa
  tmp <- plot_dat %>%
    dplyr::group_by(OTU) %>%
    dplyr::summarise(sum = sum(Abundance))
  tsums <- tmp[["sum"]]
  names(tsums) <- tmp[["OTU"]]
  tsums <- tsums[order(names(tsums))]
  
  ## Get the OTU taxa
  tmp <- subset(plot_dat, Sample == plot_dat[1, "Sample"])
  taxlist <- tmp[, taxa]
  names(taxlist) <- tmp[, "OTU"]
  
  ## Check the data match
  if (length(tsums) != length(taxlist) ||
      !all(names(tsums) %in% names(taxlist)) ||
      !all(names(taxlist) %in% names(tsums))) {
    stop("Error: tsums and taxlist do not match.\n")
  }
  taxlist <- taxlist[names(tsums)]
  
  ## Keep only the top taxa and add all other taxa into an "other" class
  taxasum <-
    sort(tapply(tsums, taxlist, sum, na.rm = TRUE), decreasing = TRUE)
  taxasum <- taxasum[!names(taxasum) == ""]
  plot_dat[["TaxName"]] <- as.character(plot_dat[[taxa]])
  plot_dat[is.na(plot_dat[["TaxName"]]) |
             plot_dat[["TaxName"]] == "", "TaxName"] <- "Other"
  plot_dat[["TaxName"]] <-
    ifelse(plot_dat[["TaxName"]] %in% names(taxasum)[1:nclass],
           as.character(plot_dat[[taxa]]),
           "Other")
  plot_dat[["TaxName"]] <-
    factor(plot_dat[["TaxName"]], levels = c(names(taxasum)[1:nclass], "Other"))
  
  ## Combine counts at the taxa level
  plot_dat <- plot_dat %>%
    dplyr::group_by(UIN, Group, TaxName) %>%
    dplyr::summarise(Abundance = sum(Abundance), .groups = "drop") %>%
    dplyr::filter(TaxName != "Other")
  
  ## Combine at the Group level
  plot_dat_SE <- plot_dat %>%
    dplyr::group_by(Group, TaxName) %>%
    dplyr::summarise(
      average = ifelse(
        test_type == "parametric",
        mean(Abundance),
        median(Abundance)
      ),
      variability = ifelse(
        test_type == "parametric",
        sd(Abundance) / sqrt(length(Abundance)),
        IQR(Abundance)
      )
    )
  
  ## Calculate p-value across all groups
  pval <-
    data.frame(TaxName       = head(levels(plot_dat_SE[["TaxName"]]),-1),
               pval          = NA)
  for (t in pval[["TaxName"]]) {
    if (test_type == "parametric") {
      pval[pval[["TaxName"]] == t, "pval"] <-
        anova(lm(Abundance ~ Group, data = subset(plot_dat, TaxName == t)))[1, "Pr(>F)"]
    } else if (test_type == "non-parametric") {
      pval[pval[["TaxName"]] == t, "pval"] <-
        kruskal.test(Abundance ~ Group, data = subset(plot_dat, TaxName == t))[["p.value"]]
    }
  }
  
  ## Calculate individual p-values vs control
  for (t in pval[["TaxName"]]) {
    for (g in unique(plot_dat[["Group"]])) {
      if (g == "Control") {
        pval[pval[["TaxName"]] == t, paste0("pval.", g)] <- 1
      } else if (test_type == "parametric") {
        pval[pval[["TaxName"]] == t, paste0("pval.", g)] <-
          t.test(
            subset(plot_dat, TaxName == t &
                     Group == g)[["Abundance"]],
            subset(plot_dat, TaxName == t &
                     Group == control)[["Abundance"]]
          )[["p.value"]]
      } else if (test_type == "non-parametric") {
        pval[pval[["TaxName"]] == t, paste0("pval.", g)] <-
          wilcox.test(
            subset(plot_dat, TaxName == t &
                     Group == g)[["Abundance"]],
            subset(plot_dat, TaxName == t &
                     Group == control)[["Abundance"]]
          )[["p.value"]]
      }
    }
  }
  
  ## Add to data frame
  plot_dat_SE[["pval"]] <- NA
  for (t in pval[["TaxName"]]) {
    for (g in unique(plot_dat[["Group"]])) {
      if (p_type == "combine") {
        if (g == control) {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <-
            pval[pval[["TaxName"]] == t, "pval"]
        } else {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <- NA
        }
      } else if (p_type == "individual") {
        if (g == control) {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <- NA
        } else {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <-
            pval[pval[["TaxName"]] == t, paste0("pval.", g)]
        }
      }
    }
  }
  
  ## Correct for multiple testing
  which_nonna <- which(!is.na(plot_dat_SE[["pval"]]))
  plot_dat_SE[["pval_adj"]] <- NA
  plot_dat_SE[which_nonna, "pval_adj"] <-
    p.adjust(as.numeric(plot_dat_SE[which_nonna,][["pval"]]),
             method = "fdr",
             n = nrow(plot_dat_SE[which_nonna,]))
  
  ## Convert to star format
  plot_dat_SE[["pval_adj_star"]] <-
    case_when(
      plot_dat_SE[["pval_adj"]] <= 0.001 ~ "***",
      plot_dat_SE[["pval_adj"]] <= 0.01  ~ "**",
      plot_dat_SE[["pval_adj"]] <= 0.05  ~ "*",
      TRUE ~ ""
    )
  
  ## Save the values
  write.table(
    plot_dat_SE,
    paste0(
      analysis_dir,
      "/tabs/plot_abundance_comparison_",
      taxa,
      ".csv"
    ),
    sep = ",",
    row.names = FALSE,
    col.names = TRUE,
    quote = FALSE
  )
  
  ## Plot the data
  
  ## Determine x-axis limits
  x_max <-
    max(plot_dat$Abundance, na.rm = TRUE) * 1.02 ## Extend by 2%
  
  ## box plots
  p2 <- plot_dat %>%
    ## remove "[]", "-", "_" from taxa names
    dplyr::mutate(TaxName = gsub("\\[|\\]|-|_", " ", TaxName)) %>%
    ## remove leading and trialing spaces!
    dplyr::mutate(TaxName = trimws(TaxName)) %>%
    ## remove instances of double spaces
    dplyr::mutate(TaxName = gsub("  ", " ", TaxName)) %>%
    ggplot(aes(
      x = Abundance,
      y = reorder(TaxName, Abundance),
      ## flip group order
      fill = factor(Group, levels = c("Intolerant", "Tolerant"))
    )) +
    geom_boxplot(
      width = 0.6,
      position = position_dodge(0.8),
      outlier.size = 1,
      outlier.alpha = 0.8
    ) +
    labs(x = "Relative Abundance",
         y = "",
         ## legend label
         fill = "Group") +
    scale_y_discrete(expand = c(0, 0.5)) +
    scale_x_continuous(
      labels = scales::label_percent(scale = 1),
      limits = c(0, x_max),
      expand = c(0, 0)
    ) +
    ## group colour for boxplot
    scale_fill_manual(values = group_cols) +
    ## group colour for jitter
    scale_color_manual(values = group_cols) +
    ## flip legend order so control above EHI
    guides(fill = guide_legend(reverse = TRUE)) +
    theme_classic() +
    theme(
      text = font,
      axis.title.x = element_text(size = 18),
      axis.text.x = element_text(size = 16),
      axis.text.y = element_markdown(size = 16, colour = "black"),
      legend.title = element_text(size = 22, face = "bold"),
      legend.text  = element_text(size = 18),
      panel.grid.major.x = element_line(colour = "darkgrey", linewidth = 0.25)
    )
  
  ## boxplot with individual data points
  p3 <- plot_dat %>%
    ## remove "[]", "-", "_" from taxa names
    dplyr::mutate(TaxName = gsub("\\[|\\]|-|_", " ", TaxName)) %>%
    ## remove leading and trialing spaces!
    dplyr::mutate(TaxName = trimws(TaxName)) %>%
    ## remove instances of double spaces
    dplyr::mutate(TaxName = gsub("  ", " ", TaxName)) %>%
    ggplot(aes(
      x = Abundance,
      y = reorder(TaxName, Abundance),
      ## flip group order
      fill = factor(Group, levels = c("Intolerant", "Tolerant"))
    )) +
    geom_jitter(
      ## colour jitter by group but flip factors to match fill global aesthetic
      aes(colour = factor(
        Group, levels = c("Intolerant", "Tolerant")
      )),
      ## set jitter width and keep dodge consistent
      position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.8),
      size = 1.2,
      alpha = 0.6,
      show.legend = FALSE,
    ) +
    geom_boxplot(width = 0.6,
                 position = position_dodge(0.8),
                 ## remove outlier dots as we have jitter points
                 outlier.shape = NA) +
    labs(x = "Relative Abundance",
         y = "",
         ## legend label
         fill = "Group") +
    scale_y_discrete(expand = c(0, 0.5)) +
    scale_x_continuous(
      labels = scales::label_percent(scale = 1),
      limits = c(0, x_max),
      expand = c(0, 0)
    ) +
    ## group colour for boxplot
    scale_fill_manual(values = group_cols) +
    ## group colour for jitter
    scale_color_manual(values = group_cols) +
    ## flip legend order so control above EHI
    guides(fill = guide_legend(reverse = TRUE)) +
    theme_classic() +
    theme(
      text = font,
      axis.title.x = element_text(size = 18),
      axis.text.x = element_text(size = 16),
      axis.text.y = element_markdown(size = 16, colour = "black"),
      legend.title = element_text(size = 22),
      legend.text  = element_text(size = 18),
      panel.grid.major.x = element_line(colour = "darkgrey", linewidth = 0.25)
    )
  
  list(BoxPlot = p2,
       JitterBoxPlot = p3)
  
} 
```

### Abundance Comparison Function: Italicised Text

The following function will plot the difference between the groups at different taxonomic levels with italicised text for taxa:

```{r Section3_plot_abundance_comparison_italised_text}
## Function for italicised taxa names
plot_abundance_comparison2 <- function (taxa,
                                        plot_dat = phyloseq_dat_filter_norm_melt,
                                        p_type = "individual",
                                        nclass = 15) {
  # taxa = "Genus"
  # plot_dat = phyloseq_dat_filter_norm_melt
  # p_type = "individual"
  
  ## Get the total abundance across all samples for each OTU for the given taxa
  tmp <- plot_dat %>%
    dplyr::group_by(OTU) %>%
    dplyr::summarise(sum = sum(Abundance))
  tsums <- tmp[["sum"]]
  names(tsums) <- tmp[["OTU"]]
  tsums <- tsums[order(names(tsums))]
  
  ## Get the OTU taxa
  tmp <- subset(plot_dat, Sample == plot_dat[1, "Sample"])
  taxlist <- tmp[, taxa]
  names(taxlist) <- tmp[, "OTU"]
  
  ## Check the data match
  if (length(tsums) != length(taxlist) ||
      !all(names(tsums) %in% names(taxlist)) ||
      !all(names(taxlist) %in% names(tsums))) {
    stop("Error: tsums and taxlist do not match.\n")
  }
  taxlist <- taxlist[names(tsums)]
  
  ## Keep only the top taxa and add all other taxa into an "other" class
  taxasum <-
    sort(tapply(tsums, taxlist, sum, na.rm = TRUE), decreasing = TRUE)
  taxasum <- taxasum[!names(taxasum) == ""]
  plot_dat[["TaxName"]] <- as.character(plot_dat[[taxa]])
  plot_dat[is.na(plot_dat[["TaxName"]]) |
             plot_dat[["TaxName"]] == "", "TaxName"] <- "Other"
  plot_dat[["TaxName"]] <-
    ifelse(plot_dat[["TaxName"]] %in% names(taxasum)[1:nclass],
           as.character(plot_dat[[taxa]]),
           "Other")
  plot_dat[["TaxName"]] <-
    factor(plot_dat[["TaxName"]], levels = c(names(taxasum)[1:nclass], "Other"))
  
  ## Combine counts at the taxa level
  plot_dat <- plot_dat %>%
    dplyr::group_by(UIN, Group, TaxName) %>%
    dplyr::summarise(Abundance = sum(Abundance), .groups = "drop") %>%
    dplyr::filter(TaxName != "Other")
  
  ## Combine at the Group level
  plot_dat_SE <- plot_dat %>%
    dplyr::group_by(Group, TaxName) %>%
    dplyr::summarise(
      average = ifelse(
        test_type == "parametric",
        mean(Abundance),
        median(Abundance)
      ),
      variability = ifelse(
        test_type == "parametric",
        sd(Abundance) / sqrt(length(Abundance)),
        IQR(Abundance)
      )
    )
  
  ## Calculate p-value across all groups
  pval <-
    data.frame(TaxName       = head(levels(plot_dat_SE[["TaxName"]]),-1),
               pval          = NA)
  for (t in pval[["TaxName"]]) {
    if (test_type == "parametric") {
      pval[pval[["TaxName"]] == t, "pval"] <-
        anova(lm(Abundance ~ Group, data = subset(plot_dat, TaxName == t)))[1, "Pr(>F)"]
    } else if (test_type == "non-parametric") {
      pval[pval[["TaxName"]] == t, "pval"] <-
        kruskal.test(Abundance ~ Group, data = subset(plot_dat, TaxName == t))[["p.value"]]
    }
  }
  
  ## Calculate individual p-values vs control
  for (t in pval[["TaxName"]]) {
    for (g in unique(plot_dat[["Group"]])) {
      if (g == "Control") {
        pval[pval[["TaxName"]] == t, paste0("pval.", g)] <- 1
      } else if (test_type == "parametric") {
        pval[pval[["TaxName"]] == t, paste0("pval.", g)] <-
          t.test(
            subset(plot_dat, TaxName == t &
                     Group == g)[["Abundance"]],
            subset(plot_dat, TaxName == t &
                     Group == control)[["Abundance"]]
          )[["p.value"]]
      } else if (test_type == "non-parametric") {
        pval[pval[["TaxName"]] == t, paste0("pval.", g)] <-
          wilcox.test(
            subset(plot_dat, TaxName == t &
                     Group == g)[["Abundance"]],
            subset(plot_dat, TaxName == t &
                     Group == control)[["Abundance"]]
          )[["p.value"]]
      }
    }
  }
  
  ## Add to data frame
  plot_dat_SE[["pval"]] <- NA
  for (t in pval[["TaxName"]]) {
    for (g in unique(plot_dat[["Group"]])) {
      if (p_type == "combine") {
        if (g == control) {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <-
            pval[pval[["TaxName"]] == t, "pval"]
        } else {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <- NA
        }
      } else if (p_type == "individual") {
        if (g == control) {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <- NA
        } else {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <-
            pval[pval[["TaxName"]] == t, paste0("pval.", g)]
        }
      }
    }
  }
  
  ## Correct for multiple testing
  which_nonna <- which(!is.na(plot_dat_SE[["pval"]]))
  plot_dat_SE[["pval_adj"]] <- NA
  plot_dat_SE[which_nonna, "pval_adj"] <-
    p.adjust(as.numeric(plot_dat_SE[which_nonna,][["pval"]]),
             method = "fdr",
             n = nrow(plot_dat_SE[which_nonna,]))
  
  ## Convert to star format
  plot_dat_SE[["pval_adj_star"]] <-
    case_when(
      plot_dat_SE[["pval_adj"]] <= 0.001 ~ "***",
      plot_dat_SE[["pval_adj"]] <= 0.01  ~ "**",
      plot_dat_SE[["pval_adj"]] <= 0.05  ~ "*",
      TRUE ~ ""
    )
  
  ## Save the values
  write.table(
    plot_dat_SE,
    paste0(
      analysis_dir,
      "/tabs/plot_abundance_comparison_",
      taxa,
      ".csv"
    ),
    sep = ",",
    row.names = FALSE,
    col.names = TRUE,
    quote = FALSE
  )
  
  ## Plot the data
  
  ## Determine x-axis limits
  x_max <-
    max(plot_dat$Abundance, na.rm = TRUE) * 1.02 ## Extend by 2%
  
  ## box plots
  p2 <- plot_dat %>%
    ## remove "[]", "-", "_" from taxa names
    dplyr::mutate(TaxName = gsub("\\[|\\]|-|_", " ", TaxName)) %>%
    ## remove leading and trialing spaces!
    dplyr::mutate(TaxName = trimws(TaxName)) %>%
    ## remove instances of double spaces
    dplyr::mutate(TaxName = gsub("  ", " ", TaxName)) %>%
    dplyr::mutate(
      FormattedTaxName = ifelse(TaxName != "Other", paste("*", TaxName, "*", sep = ""), TaxName),
      FormattedTaxName = ifelse(
        FormattedTaxName == "*Eubacterium hallii group*",
        "*Eubacterium<br>hallii group*",
        FormattedTaxName
      )
    ) %>%
    ggplot(aes(
      x = Abundance,
      y = reorder(FormattedTaxName, Abundance),
      ## flip group order
      fill = factor(Group, levels = c("Intolerant", "Tolerant"))
    )) +
    geom_boxplot(
      width = 0.6,
      position = position_dodge(0.8),
      outlier.size = 1,
      outlier.alpha = 0.8
    ) +
    labs(x = "Relative Abundance",
         y = "",
         ## legend label
         fill = "Group") +
    scale_y_discrete(expand = c(0, 0.5)) +
    scale_x_continuous(
      labels = scales::label_percent(scale = 1),
      limits = c(0, x_max),
      expand = c(0, 0)
    ) +
    ## group colour for boxplot
    scale_fill_manual(values = group_cols) +
    ## group colour for jitter
    scale_color_manual(values = group_cols) +
    ## flip legend order so control above EHI
    guides(fill = guide_legend(reverse = TRUE)) +
    theme_classic() +
    theme(
      text = font,
      axis.title.x = element_text(size = 18),
      axis.text.x = element_text(size = 16),
      axis.text.y = element_markdown(size = 16, colour = "black"),
      legend.title = element_text(size = 22, face = "bold"),
      legend.text  = element_text(size = 18),
      panel.grid.major.x = element_line(colour = "darkgrey", linewidth = 0.25)
    )
  
  ## boxplot with individual data points
  p3 <- plot_dat %>%
    ## remove "[]", "-", "_" from taxa names
    dplyr::mutate(TaxName = gsub("\\[|\\]|-|_", " ", TaxName)) %>%
    ## remove leading and trialing spaces!
    dplyr::mutate(TaxName = trimws(TaxName)) %>%
    ## remove instances of double spaces
    dplyr::mutate(TaxName = gsub("  ", " ", TaxName)) %>%
    dplyr::mutate(
      FormattedTaxName = ifelse(TaxName != "Other", paste("*", TaxName, "*", sep = ""), TaxName),
      FormattedTaxName = ifelse(
        FormattedTaxName == "*Eubacterium hallii group*",
        "*Eubacterium<br>hallii group*",
        FormattedTaxName
      )
    ) %>%
    ggplot(aes(
      x = Abundance,
      y = reorder(FormattedTaxName, Abundance),
      ## flip group order
      fill = factor(Group, levels = c("Intolerant", "Tolerant"))
    )) +
    geom_jitter(
      ## colour jitter by group but flip factors to match fill global aesthetic
      aes(colour = factor(
        Group, levels = c("Intolerant", "Tolerant")
      )),
      ## set jitter width and keep dodge consistent
      position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.8),
      size = 1.2,
      alpha = 0.6,
      show.legend = FALSE,
    ) +
    geom_boxplot(width = 0.6,
                 position = position_dodge(0.8),
                 ## remove outlier dots as we have jitter points
                 outlier.shape = NA) +
    labs(x = "Relative Abundance",
         y = "",
         ## legend label
         fill = "Group") +
    scale_y_discrete(expand = c(0, 0.5)) +
    scale_x_continuous(
      labels = scales::label_percent(scale = 1),
      limits = c(0, x_max),
      expand = c(0, 0)
    ) +
    ## group colour for boxplot
    scale_fill_manual(values = group_cols) +
    ## group colour for jitter
    scale_color_manual(values = group_cols) +
    ## flip legend order so control above EHI
    guides(fill = guide_legend(reverse = TRUE)) +
    theme_classic() +
    theme(
      text = font,
      axis.title.x = element_text(size = 18),
      axis.text.x = element_text(size = 16),
      axis.text.y = element_markdown(size = 16, colour = "black"),
      legend.title = element_text(size = 22),
      legend.text  = element_text(size = 18),
      panel.grid.major.x = element_line(colour = "darkgrey", linewidth = 0.25)
    )
  
  list(BoxPlot = p2,
       JitterBoxPlot = p3)
  
} 
```
  
### Phylum
``` {r Section3_plot_abundance_comparison_phylum, fig.height = 5, fig.width = 7}
plot_abundance_comparison("Phylum", nclass = 7)

phylum_abund_outcome <- plot_abundance_comparison("Phylum", nclass = 7)
phylum_abund_outcome_p <- phylum_abund_outcome$JitterBoxPlot
```

### Class
``` {r Section3_plot_abundance_comparison_class, fig.height = 5, fig.width = 7}
plot_abundance_comparison("Class")
```

### Order
``` {r Section3_plot_abundance_comparison_order, fig.height = 5, fig.width = 7}
plot_abundance_comparison("Order")
```

### Family
``` {r Section3_plot_abundance_comparison_family, fig.height = 5, fig.width = 7}
plot_abundance_comparison2("Family")
```

### Genus
``` {r Section3_plot_abundance_comparison_genus, fig.height = 5, fig.width = 7}
plot_abundance_comparison2("Genus")

genus_abund_outcome <- plot_abundance_comparison2("Genus")
genus_abund_outcome_p <- genus_abund_outcome$JitterBoxPlot
```

# Section 4: Differential Abundance Analysis

## DESeq2

In this section, we will use the [DESeq2](https://bioconductor.org/packages/release/bioc/html/DESeq2.html) package in R to identify differentially abundant bacterial species at the OTU level between the conditions. Below are a number of functions that will generate results and plots for the various Group comparisons that will be made.

### Setup

``` {r Section4_DESeq2_analysis_setup}
## Perform the differential expression analysis
## Run DESeq2 on non-rarefied feature table
deseq_dat <- phyloseq_dat_deseq
## plot the abundance non-rarefied feature table
deseq_plot <- phyloseq_dat_norm_melt 
```

### Model

``` {r Section4_DESeq2_analysis_dds}
## create DESeq2 object from phyloseq object
dds <- phyloseq_to_deseq2(deseq_dat, design = ~ Group)

## Pre-filtering
smallestGroupSize <- 3
keep <- rowSums(counts(dds) >= 10) >= smallestGroupSize
dds <- dds[keep,]

## estimation of size factors was set to use poscounts, which calculates a modified relative log expression that helps account for features missing in at least one sample
dds <- DESeq(dds, test = "Wald", fitType = "local", sfType = "poscounts")

## Plot deseq2 object
plotDispEsts(dds)
## coefficients DESeq estimated:
resultsNames(dds)

## results
res_null <- results(dds)
## summary 
summary(res_null)
```

### Function

``` {r Section4_DESeq2_analysis_function}
## Function to take the results from a specific contrast from DESeq2 and return a significance table
signif_OTU <- function (diff_OTU) {
  diff_OTU <-
    subset(diff_OTU,!is.na(padj) &
             padj <= 0.05 & abs(log2FoldChange) >= 1)
  if (nrow(diff_OTU) == 0)
    return(NULL)
  diff_OTU <-
    cbind(as(diff_OTU, "data.frame"), as(tax_table(deseq_dat)[rownames(diff_OTU),], "matrix"))
  diff_OTU <-
    diff_OTU[order(diff_OTU[["padj"]], decreasing = FALSE), ]
  diff_OTU[["OTU"]] <- rownames(diff_OTU)
  diff_OTU[["OTU_num"]] <- 1:nrow(diff_OTU)
  return(diff_OTU)
}

## Function to plot a specific OTU
plot_OTU <- function (otu) {
  ggplot(subset(deseq_plot, OTU == otu),
         aes(x = Group, y = Abundance, fill = Group)) +
    geom_boxplot() +
    ggtitle(paste(c(otu, as.character(unlist(
      subset(deseq_plot, OTU == otu)[1, c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus")]
    ))), collapse = "; ")) +
    labs(x = "", y = "Relative Abundance") +
    scale_fill_manual(values = group_cols) +
    scale_y_continuous(labels = scales::label_percent(scale = 1)) +
    theme(
      text = font,
      plot.title = element_text(size = 8),
      axis.title   = element_text(size = 18),
      axis.text.y  = element_text(size = 16),
      axis.text.x  = element_text(
        angle = 90,
        hjust = 1,
        size = 24
      ),
      legend.text  = element_text(size = 22),
      legend.title = element_text(size = 22, face = "bold")
    )
}

## Function to process a specific comparison
diff_out <- function (i, j, plot = TRUE) {
  ## Get the significant OTUs
  if (i == j) {
    cat ("Error: ", i, " and ", j, " must be different\n")
    return (NULL)
  }
  all_OTU <- results(dds, contrast = c("Group", i, j))
  diff_OTU <- signif_OTU(all_OTU)
  if (is.null(diff_OTU)) {
    cat ("Error: No significant differences identified for ",
         i,
         " vs ",
         j,
         "\n")
    return (list(
      "All"    = all_OTU,
      "Signif" = NULL,
      "Up"     = NULL,
      "Down"   = NULL
    ))
  }
  
  ## Save the results
  write.table(
    diff_OTU,
    paste0(
      analysis_dir,
      "/tabs/DESeq2_",
      i,
      "_vs_",
      j,
      "_log2foldchange.csv"
    ),
    row.names = FALSE,
    col.names = TRUE,
    sep = ",",
    quote = FALSE
  )
  
  ## Plot the data for these OTUs
  if (plot) {
    outdir <-
      paste0(analysis_dir, "/figs/SIGNIF_OTUs2_", i, "_vs_", j, "/")
    if (file.exists(outdir)) {
      unlink(outdir, recursive = TRUE)
      dir.create(outdir)
    }
    for (k in diff_OTU[["OTU_num"]]) {
      otu <- rownames(diff_OTU)[k]
      ggsave(
        filename = paste0(outdir, "OTU_", k, "_", otu, ".png"),
        plot = plot_OTU(otu),
        height = 6,
        width = 8,
        unit = "in",
        dpi = 300
      )
    }
  }
  
  ## Return the results
  return (list(
    "All"    = all_OTU,
    "Signif" = diff_OTU,
    "Up"     = subset(diff_OTU, log2FoldChange > 0),
    "Down"   = subset(diff_OTU, log2FoldChange < 0)
  ))
}

## Function to plot a volcano plot for the results using ggplot2
plot_volcano <- function(x) {
  if (length(x) == 0)
    return(NULL)
  
  # Create a data frame for plotting
  plot_dat <- x[["All"]]
  plot_col <- rep("black", nrow(plot_dat))
  which_down <- which(rownames(plot_dat) %in% rownames(x[["Down"]]))
  which_up <- which(rownames(plot_dat) %in% rownames(x[["Up"]]))
  plot_col[which_down] <- "blue"
  plot_col[which_up] <- "red"
  plot_dat_df <- data.frame(
    log2FoldChange = plot_dat$log2FoldChange,
    padj = plot_dat$padj,
    plot_col = factor(plot_col)
  )
  
  # Create the volcano plot using ggplot2
  p <-
    ggplot(data = plot_dat_df, aes(
      x = log2FoldChange,
      y = -log10(padj),
      color = plot_col
    )) +
    geom_point(size = 3,
               #shape = 1,
               alpha = 0.3) +
    scale_color_manual(values = c("black", "blue", "red")) +
    labs(x = "log2(Fold Change)", y = "-log10(P Value)") +
    # scale_x_continuous(breaks = seq(-6, 6, 2)) +
    theme_classic() +
    theme(
      text = font,
      legend.position = "none",
      plot.title = element_text(size = 12),
      axis.title = element_text(size = 14),
      axis.text = element_text(size = 16),
      legend.text = element_text(size = 12),
      legend.title = element_text(size = 12, face = "bold")
    ) +
    geom_hline(
      yintercept = -log10(0.05),
      linetype = "dashed",
      color = "black"
    ) +
    geom_vline(
      xintercept = log2(2),
      linetype = "dashed",
      color = "black"
    ) +
    geom_vline(
      xintercept = -log2(2),
      linetype = "dashed",
      color = "black"
    )
  
  return(p)
}

## Function to get the top hits from the DESeq2 analysis
get_top_hits <- function(x, n = 10) {
  if (length(x) == 0)
    return()
  
  # Get the top up and downregulated hits
  positive_df <-
    head(x[["Up"]][order(x[["Up"]][["log2FoldChange"]], decreasing = TRUE),], min(n, nrow(x[["Up"]])))
  negative_df <-
    tail(x[["Down"]][order(x[["Down"]][["log2FoldChange"]], decreasing = TRUE),], min(n, nrow(x[["Down"]])))
  df <- rbind.data.frame(positive_df, negative_df)
  df[["Direction"]] <-
    c(rep("Increase", min(n, nrow(x[["Up"]]))), rep("Decrease", min(n, nrow(x[["Down"]]))))
  
  # Set the sample names from the "Genus" column
  # df[["Taxa"]] <- factor(paste(df[["OTU_num"]], df[["Genus"]], sep = ": "),
  #                                levels = paste(df[["OTU_num"]], df[["Genus"]], sep = ": "))
  
  df[["Taxa"]] <-
    factor(paste(otu_or_asv, df[["OTU_num"]], ": ", df[["Genus"]], sep = ""),
           levels = paste(otu_or_asv, df[["OTU_num"]], ": ", df[["Genus"]], sep = ""))
  
  # Create a bar plot using the log2FoldChange values and updated sample names, with colored bars
  df %>%
    ggplot(aes(x = Taxa, y = log2FoldChange, fill = Direction)) +
    geom_bar(stat = 'identity') +
    coord_flip() +
    scale_fill_manual(values = c(
      'Increase' = "#117744",
      'Decrease' = "#DC3220"
    )) +
    ylab("log2 (Fold Change)") + xlab("") +
    # scale_y_continuous(breaks = seq(-6, 6, 2)) +
    theme_classic() +
    theme(
      text = font,
      axis.title   = element_text(size = 14),
      axis.text.y  = element_text(size = 16, face = "bold"),
      axis.text.x  = element_text(size = 14),
      legend.position = "none",
      panel.grid.major.x = element_line(colour = "lightgrey", size = 0.25)
    )
}
```

### Heat Tolerant vs Heat Intolerant
The following results show the differentially abundant taxa between the Heat Tolerant and Heat Intolerant samples. 

``` {r Section4_DESeq2_analysis_intolerant_vs_tolerant}
DE_PvF <- diff_out("Intolerant", "Tolerant", plot = FALSE)
cat("Total number of differentially abundant OTUs: ", nrow(DE_PvF[["Signif"]]))
cat("Upregulated number of differentially abundant OTUs: ", nrow(DE_PvF[["Up"]]))
cat("Downregulated number of differentially abundant OTUs: ", nrow(DE_PvF[["Down"]]))
```

Below is a volcano plot showing the overall differential abundance results for this comparison.

``` {r Section4_DESeq2_analysis_intolerant_vs_tolerant_volcano, fig.height = 5, fig.width = 7}
volcano_p <- plot_volcano(DE_PvF)
volcano_p
```

## ALDEx2

### Setup

```{r Section4_ALDEx2_setup}
## create ASV_table from phyloseq object
ASV_table = otu_table(phyloseq_dat_deseq)

## create groupings variable from phyloseq object
groupings = phyloseq_dat_deseq %>%
  sample_data() %>%
  .$Group %>%
  ## make groupings a character vector! 
  as.character()

ncol(ASV_table)
length(groupings)
```

### Model

```{r Section4_ALDEx2_model}
results <- aldex(reads = ASV_table, 
                 conditions = groupings, 
                 mc.samples = 128, 
                 test="t", 
                 effect=TRUE,
                 include.sample.summary = TRUE, 
                 CI = TRUE,
                 verbose = TRUE, 
                 denom="all")
```

### Results

```{r Section4_ALDEx2_results}
## Expected p-value of Wilcoxon rank test
results %>%
  dplyr::filter(we.eBH < 0.05)

## Expected Benjamini-Hochberg corrected p-value of Wilcoxon test
results %>%
  dplyr::filter(wi.eBH < 0.05)
```

## ANCOMBC

### Setup

https://bioconductor.org/packages/devel/bioc/vignettes/ANCOMBC/inst/doc/ANCOM.html

```{r Section4_ANCOMBC_setup}
## object to use
phyloseq_dat_deseq 
## make a copy for ancom analysis
ancom_phyloseq = phyloseq_dat_deseq
## Set Group variable as character string 
sample_data(ancom_phyloseq)$Group <- as.character(sample_data(ancom_phyloseq)$Group)
## Check that levels are now NULL
levels(sample_data(ancom_phyloseq)$Group)
## Make sure names are still "Tolerant" and "Intolerant"
unique(sample_data(ancom_phyloseq)$Group)
```

### Model

```{r Section4_ANCOMBC_model}
## use ANCOMBC with phyloseq object
out = ancombc(data = NULL, 
              assay_name = NULL,
              tax_level = NULL, 
              phyloseq = phyloseq_dat_deseq,
              formula = "Group",
              p_adj_method = "BH", 
              prv_cut = 0.10, 
              lib_cut = min_otus,
              group = "Group", 
              struc_zero = TRUE, 
              neg_lb = TRUE,
              tol = 1e-5, 
              max_iter = 100, 
              conserve = FALSE,
              alpha = 0.05, 
              global = TRUE, 
              n_cl = 1, 
              verbose = TRUE)
```

### Results

```{r Section4_ANCOMBC_results}
## Result from the ANCOM-BC log-linear model to determine taxa that are differentially abundant according to the covariate of interest. It contains: 1) log fold changes; 2) standard errors; 3) test statistics; 4) p-values; 5) adjusted p-values; 6) indicators whether the taxon is differentially abundant (TRUE) or not (FALSE).
res = out$res

## Set Colnames
col_names = c("Taxon", "Intercept", "Group")

## 1) log fold changes
res_lfc = res$lfc
colnames(res_lfc) = col_names
#res_lfc
## 2) standard errors
res_se = res$se
colnames(res_se) = col_names
#res_se
## 3) test statistics
res_w = res$W
colnames(res_w) = col_names
#res_w
## 4) p-values
res_p = res$p_val
colnames(res_p) = col_names
#res_p
## 5) adjusted p-values
res_padj = res$q_val
colnames(res_padj) = col_names
#res_padj
## 6) indicators whether the taxon is differentially abundant (TRUE) or not (FALSE).
res_da = res$diff_abn
colnames(res_da) = col_names
#res_da
## check if we have TRUE
unique(res_da$Group)
```

```{r Section4_ANCOM_save_results}
## Create list of ANCOM results
ancom_res_list <- list(
  "Log_Fold_Change" = res_lfc,
  "Standard_Errors" = res_se,
  "Test_Statistics" = res_w,
  "p_value" = res_p,
  "BH_Adjusted_p_value" = res_padj,
  "Dif_Abun_T_or_F" = res_da
)

## Save ANCOM results into a single workbook
openxlsx::write.xlsx(ancom_res_list,
                     file = paste0(analysis_dir, "/tabs/ANCOM_results.xlsx"))
```

## Compare Methods

```{r Section4_compare}
## DESeq2 results to use
deseq_res = DE_PvF$Signif
## set rownames as a new colmun called Taxon
deseq_res = rownames_to_column(deseq_res, var = "Taxon")
## count number of sig ASVs using nrow()
nrow(deseq_res)
## Select only Taxon column
deseq_res = deseq_res %>%
  dplyr::select(Taxon)

## ALDEx2 results to use
aldex2_res = results
## set rownames as a new colmun called Taxon
aldex2_res = rownames_to_column(aldex2_res, var = "Taxon")
## Filter for only p adjust < 0.05
aldex2_res = results %>%
  dplyr::filter(wi.eBH < 0.05)
## count number of sig ASVs using nrow()
nrow(aldex2_res)
## As we have 0 hits, we will create an empty vector for venn diagram
aldex2_hits = character(nrow(aldex2_res))

## ANCOMBC results, filter for only p adjust < 0.05
ancombc_res = res_padj %>%
  dplyr::filter(Group < 0.05)
## count number of sig ASVs using nrow()
nrow(ancombc_res)
## Select only Taxon column
ancombc_res = ancombc_res %>%
  dplyr::select(Taxon)

## Prepare the a list for venn diagram plotting.
da_list <- list(
  DESeq2 = deseq_res$Taxon,
  ALDEx2 = aldex2_hits,  
  "ANCOM-BC" = ancombc_res$Taxon
)
```

### Venn Diagram

```{r Section4_venn_diagram, fig.height = 8, fig.width = 8}
## Change font, code from https://stackoverflow.com/questions/6699950/r-geom-venn-how-to-change-font
venn_font <- function(p, font)
{

  grep_grob <- function(gt, lab){
    which(sapply(gt, function(x) grepl(lab, x$name)))
  }
  
  p2 <- ggplot_gtable(ggplot_build(p))
  mygrobs <- p2$grobs
  panel_grob <- mygrobs[[grep_grob(mygrobs, "panel")]]
  venn_grob <- panel_grob$children[[grep_grob(panel_grob$children, "venn")]]
  text_grob <- venn_grob$children[grep_grob(venn_grob$children, "text")]
  text_grob <- do.call(grid::gList, 
                       lapply(text_grob, 
                              function(x) {x$gp$fontfamily <- font; 
                                           x}))
  venn_grob$children[grep_grob(venn_grob$children, "text")] <- text_grob
  panel_grob$children[[grep_grob(panel_grob$children, "venn")]] <- venn_grob
  mygrobs[[grep_grob(mygrobs, "panel")]] <- panel_grob
  p2$grobs <- mygrobs
  grid::grid.newpage()
  grid::grid.draw(p2)
}

## Combine all taxa into a single vector and make it unique
venn_all_taxa <- unique(unlist(da_list))

## Create a data frame where each row is a unique taxa
venn_df <- data.frame(Taxon = venn_all_taxa)

## For each method, create a column in taxa_df indicating presence or absence of each taxon
for(method in names(da_list)) {
  venn_df[[method]] <- venn_df$Taxon %in% da_list[[method]]
}

## view
venn_df

## create venn diagram
venn_p <- ggplot(venn_df) +
  geom_venn(
    aes(A = DESeq2, B = `ALDEx2`, C = `ANCOM-BC`),
    fill_color = c("#E69F00", "#56B4E9", "green"),
    text_size = 5
  ) +
  coord_fixed() +
  theme_void()

venn_p

## use custom function to set font as serif
venn_font(venn_p, font = "serif")
```

# Section 6: Positive Control Analysis
This data set contained samples generated from the [Zymobiomics Microbial Community Standard](https://zymoresearch.eu/collections/zymobiomics-microbial-community-standards/products/zymobiomics-microbial-community-standard). This is used to assess bias and errors in the extraction methods and identify potential contamination across the data set. 

## Theoretical distribution
These samples have a theoretical Composition Based on Genomic DNA which should be represented in the output distributions following data processing alongside the other samples in the data set. This distribution is as follows:

Listeria monocytogenes - 12%, Pseudomonas aeruginosa - 12%, Bacillus subtilis - 12%, Escherichia coli - 12%, Salmonella enterica - 12%, Lactobacillus fermentum - 12%, Enterococcus faecalis - 12%, Staphylococcus aureus - 12%, Saccharomyces cerevisiae - 2%, and Cryptococcus neoformans - 2%

Note that only the bacterial species (all excluding Saccharomyces cerevisiae and Cryptococcus neoformans) will be identified through 16S rRNA amplicon based sequencing. The theoretical distribution for only the bacterial species is shown below:

``` {r Section6_PC_barplot_theoretical_abundance_stacked, fig.height = 10, fig.width = 6.5}
theo <- data.frame(species = c("*Listeria monocytogenes*", "*Pseudomonas aeruginosa*", 
                               "*Bacillus subtilis*", "*Escherichia coli*", 
                               "*Salmonella enterica*", "*Lactobacillus fermentum*", 
                               "*Enterococcus faecalis*", "*Staphylococcus aureus*"),
                   percent = rep(12.5, 8),
                   rep     = "1",
                   group = rep("Positive Control<br>Theoretical Distribution", 8))

theo_cols = c("*Bacillus subtilis*" = "#771122",
              "*Escherichia coli*" = "#114477",
              "*Lactobacillus fermentum*" = "#77CCCC",
              "*Staphylococcus aureus*" = "#AA4488",
              "*Salmonella enterica*" = "#DDDD77",
              "*Listeria monocytogenes*" = "#AA7744",
              "*Enterococcus faecalis*" = "#777711",
              "*Pseudomonas aeruginosa*" = "#AAAA44"
              )

theo_plot <- ggplot(theo, aes(x = rep, 
                 y = percent, 
                 fill = species)) + 
  geom_bar(stat = "identity") +
  scale_y_continuous(labels = scales::label_percent(scale = 1),
                     expand = expansion(add = c(0, 1))) +
  scale_fill_manual(values = theo_cols) +
  labs(x = "", 
       y = "Relative Abundance") + 
  facet_wrap(~group) +
  theme_bw() +
  theme(
    text = font,
    axis.title   = element_text(size = 18),
    axis.text.y  = element_text(size = 16),
    axis.text.x  = element_blank(),
    legend.text  = element_markdown(size = 16),
    legend.title = element_blank(),
    strip.text.x = element_markdown(size = 16, face = "bold")) +
  guides(fill = guide_legend(ncol = 1))

theo_plot
```

The total number of taxa (in thousands) for the PC samples is shown below.

``` {r Section6_QC_OTU_count_PC, fig.height = 6, fig.width = 10}
## print counts
data.frame(sample_data(phyloseq_dat_PC)) %>%
  dplyr::mutate(count = sample_sums(phyloseq_dat_PC)) %>%
  dplyr::mutate(Group = str_replace(Group, "PositiveControl", "Positive Control")) %>%
  dplyr::select(sample_code, Rep, count)

## plot counts
data.frame(sample_data(phyloseq_dat_PC)) %>%
  dplyr::mutate(count = sample_sums(phyloseq_dat_PC)) %>%
  dplyr::mutate(Group = str_replace(Group, "PositiveControl", "Positive Control")) %>%
  ggplot(aes(x = Rep, y = count / 1000, fill = Rep)) +
  geom_bar(stat = "identity",
           position = "dodge",
           color = "black") +
  xlab("") + 
  ylab(paste0("Total Number ", otu_or_asv, "s (x1000)")) +
  facet_wrap(. ~ Group, scales = "free_x") +
  geom_hline(yintercept = min_otus / 1000,
             linetype = 2,
             linewidth = 1.2) +
  scale_fill_manual(values = custom_col16) +
  theme_bw() +
  theme(
    text = font,
    axis.title   = element_text(size = 18),
    axis.text.y  = element_text(size = 12),
    axis.text.x  = element_text(
      size = 14,
      angle = 315,
      face = "bold",
      hjust = 0
    ),
    legend.position = "none",
    strip.text.x = element_text(size = 20, face = "bold"),
    strip.text.y = element_text(size = 16, face = "bold")
  )
```

Then we filter to remove the samples with too few taxa, rarefy to the depth of the lowest sample, and normalise to percentage abundance as we did above.

``` {r Section6_PC_analysis_setup}
## Filter to remove samples with fewer than min_otus
phyloseq_dat_PC_filter <- prune_samples(sample_sums(phyloseq_dat_PC) >= min_otus, phyloseq_dat_PC)

## Rarefy the data to the minimum otu depth
set.seed(0)
rarefy_otu_depth <- min(sample_sums(phyloseq_dat_PC_filter))
cat("Rarefication of data to", rarefy_otu_depth, "OTUs:\n")
phyloseq_dat_PC_filter <- rarefy_even_depth(phyloseq_dat_PC_filter, rarefy_otu_depth, verbose = FALSE)

## Normalise the data
phyloseq_dat_PC_filter_norm <- transform_sample_counts(phyloseq_dat_PC_filter, function (x) 100* ((x+1) / sum(x+1)))

## Create data for barplots
phyloseq_dat_PC_filter_norm_melt <- psmelt(phyloseq_dat_PC_filter_norm) %>%
  dplyr::mutate(Genus = ifelse (Genus == "Limosilactobacillus", "Lactobacillus", Genus)) %>% 
  dplyr::mutate(Genus = ifelse (Genus == "Escherichia-Shigella", "Escherichia", Genus)) %>%
  dplyr::mutate(Group = str_replace(Group, "PositiveControl", "Positive Control"))
```

## Chi-squared

Below is a chi-squared based comparison between the PC replicates against the expected distribution:

``` {r Section_6_PC_comparison_with_expected}
## Get the theoretical distribution
PC_comp <- theo %>%
  ## remove "*" from names
  dplyr::mutate(species = gsub("\\*", "", species))
PC_comp[["Genus"]] <- gsub(" .*$", "", PC_comp[["species"]])
PC_comp <- PC_comp[, c("Genus", "percent")]
names(PC_comp)[2] <- "Zymo"

## Get the abundance for all of the species in the positive control
PC_abun <- phyloseq_dat_PC_filter_norm_melt %>%
  ## Change where NA is present for Genus column to Salmonella
  dplyr::mutate(Genus = ifelse(!is.na(Family) & Family == "Enterobacteriaceae" & is.na(Genus), "Salmonella", Genus)) %>%
  ## Make uncultured NA
  dplyr::mutate(Genus = ifelse(is.na(Genus) | Genus == "uncultured", NA, Genus)) %>%
  ## Set levels
  dplyr::mutate(Genus = factor(Genus, levels = unique(Genus))) %>%
  dplyr::group_by(Genus, Sample) %>%
  dplyr::summarise(Abundance = sum(Abundance), .groups = "drop") %>%
  dplyr::filter(Genus %in% PC_comp[["Genus"]]) %>%
  tidyr::spread(key = Sample, value = Abundance)

PC_abun

## Combine and save
PC_abun_merge <- merge(PC_comp, PC_abun, by = "Genus")
write.table(
  PC_abun_merge,
  paste0(analysis_dir, "/tabs/PC_abundance_comparison.csv"),
  sep = ",",
  row.names = FALSE,
  col.names = TRUE,
  quote = FALSE
)

## Chi-squared test: all data including theo
csqtest_theo <- chisq.test(PC_abun_merge[, -1])
csqtest_theo

## Just compare replicates to each other
csqtest <- chisq.test(PC_abun_merge[, -c(1:2)])
csqtest

## Replicate Theo 5 times intotal to balance comparison to 5 theo vs 5 positive control reps
chisq.test(PC_abun_merge[, c(rep(2,5), 3:7)])

chisq.test(PC_abun_merge[, c(2,3)])
chisq.test(PC_abun_merge[, c(2,4)])
chisq.test(PC_abun_merge[, c(2,5)])
chisq.test(PC_abun_merge[, c(2,6)])
chisq.test(PC_abun_merge[, c(2,7)])
```

## Plot

The distribution of taxa at the Genus level is shown below for the PC samples:

``` {r Section6_PC_barplot_top_genus_normalised_abundance_stacked, fig.height = 8, fig.width = 6.5}
## create new dataframe to use for positive controls
pc_df <- phyloseq_dat_PC_filter_norm_melt
## Change where NA is present for Genus column to Salmonella
pc_df[!is.na(pc_df[["Family"]]) & pc_df[["Family"]] == "Enterobacteriaceae" & is.na(pc_df[["Genus"]]), "Genus"] <- "Salmonella"
## make uncultured NA
pc_df[is.na(pc_df[["Genus"]]) | pc_df[["Genus"]] == "uncultured", "Genus"] <- NA
## set levels
pc_df$Genus <- factor(pc_df$Genus, levels = unique(pc_df$Genus))
## now plot barplot
PC_plot_df <- plot_barplot2("Genus", pc_df, "Positive_Control")
## show plot
PC_plot_df[[1]] 
## Save plot for additional plotting
PC_plot <- PC_plot_df[[1]] 

## chi-squared stats
stat <- round(csqtest$statistic,3)
df <- csqtest$parameter
p_value <- round(csqtest$p.value, 3)

## title
title <- bquote("Chi-Squared: " ~ {X^2}["(" * .(df) * ")"] ~ "=" ~ .(stat) * ", " ~ italic(p) ~ "=" ~ .(p_value))
caption <- "Chi-Squared test comparing all replicates."

PC_plot <- PC_plot +
  scale_x_discrete(labels = c("Rep 1", "Rep 2", "Rep 3", "Rep 4", "Rep 5")) +
  theme_classic() +
  labs(title = title,
       caption = caption) +
  facet_wrap(~Group) +
  theme(
    text = font,
    title = element_text(size = 16),
    plot.caption = element_text(size = 12),
    plot.caption.position = "plot",
    axis.title   = element_text(size = 24),
    axis.text.y  = element_text(size = 16),
    axis.text.x  = element_text(size = 22, angle = 90, vjust = 0.5),
    legend.text  = element_markdown(size = 14),
    legend.title = element_blank(),
    strip.text.x = element_text(size=26, face = "bold")) 

## view
PC_plot

## PC plot with theo, and chi-squared test stats including theo 
## chi-squared stats
stat_theo <- round(csqtest_theo$statistic,3)
df_theo <- csqtest_theo$parameter
p_value_theo <- round(csqtest_theo$p.value, 3)

## title
title <- bquote("Chi-Squared: " ~ {X^2}["(" * .(df_theo) * ")"] ~ "=" ~ .(stat_theo) * ", " ~ italic(p) ~ "=" ~ .(p_value_theo))
caption <- "Chi-Squared test comparing all replicates and theoretical distribution."

PC_plot_theo <- theo_plot +
  theme_classic() +
  labs(title = title,
       caption = caption) +
  theme(
    text = font,
    title = element_text(size = 16),
    plot.caption = element_text(size = 12),
    plot.caption.position = "plot",
    axis.title   = element_text(size = 24),
    axis.text.y  = element_text(size = 16),
    axis.text.x  = element_blank(),
    legend.text  = element_markdown(size = 14),
    legend.title = element_blank(),
    strip.text.x = element_markdown(size=16, face = "bold")) 

PC_plot_theo
```

# Section 7: Negative Control Analysis

## Taxa count
The total number of taxa for the NFW samples is shown below. This should be as low as possible for all replicates 

``` {r Section7_QC_OTU_count_NFW, fig.height = 8, fig.width = 6}
NC_plot <- data.frame(sample_data(phyloseq_dat_NFW)) %>%
  dplyr::mutate(count = sample_sums(phyloseq_dat_NFW)) %>%
  dplyr::mutate(Group = str_replace(Group, "NegativeControl", "Negative Control")) %>%
  ggplot(aes(x = Rep, y = count, fill = Rep)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  xlab("") + 
  #ylab(paste0("Total Number ", otu_or_asv, "s")) +
  ylab("Read Count") +
  facet_wrap(.~Group, scales = "free_x") +
  #scale_x_discrete(expand=c(0,0)) +
  scale_y_continuous(expand = expansion(add = c(0, 1))) +
  theme_classic() +
    theme(
      text = font,
      axis.title   = element_text(size = 24),
      axis.text.y  = element_text(size = 16),
      axis.text.x  = element_text(size = 22, angle = 90, vjust = 0.5),
      legend.position = "none",
      strip.text.x = element_text(size=26, face = "bold")
      )

## view
NC_plot

## print rownames (otu id) with counts greater than 1
otu_table(phyloseq_dat_NFW)[apply(otu_table(phyloseq_dat_NFW), 1, max) > 1, ]

## check tax table 
tax_table(phyloseq_dat_NFW)["39f7a3642463f94ed9c07eea3caa999e",]
tax_table(phyloseq_dat_NFW)["a4ab770f8e7c36e376f7c2e36a4611cd",]

## plot caption
caption <- "Rep 1 contains one unique ASV which is \"Unassigned\".<br>Rep 4 contains one unique ASV \"Bacteria\" which is NA at each taxonomic level."

## plot
NC_plot_caption <- NC_plot +
  labs(caption = caption) +
  theme(plot.caption = element_markdown(size = 12))

NC_plot_caption
```

Then we filter normalise to percentage abundance as we did above (note that we are not filtering or rarefying these sample, as we want to observe the negative influences on the data).

``` {r Section7_NFW_analysis_setup}
## Normalise the data
phyloseq_dat_NFW_filter_norm <- transform_sample_counts(phyloseq_dat_NFW, function (x) 100* ((x+1) / sum(x+1)))

## Create data for barplots
phyloseq_dat_NFW_filter_norm_melt <- psmelt(phyloseq_dat_NFW_filter_norm) %>%
   dplyr::mutate(Group = str_replace(Group, "NegativeControl", "Negative Control"))
```

The distribution of taxa at the Genus level is shown below for the NFW samples (note that these data are NOT filtered, as this would remove all samples - so this is based on very low levels of reads). Note that this plot may be influenced by the fact that all taxa had one added to them when normalising to avoid divide by zero errors, so if the read count is low then this plot can be ignored.

``` {r Section7_NFW_barplot_top_genus_normalised_abundance_stacked, fig.height = 8, fig.width = 7}
genus_dat_NFW <- plot_barplot2("Genus",
                               phyloseq_dat_NFW_filter_norm_melt,
                               "negative_control")
genus_dat_NFW[[1]]
```

# Section 8: Figures

## Microbiome Panel

```{r Section8_microbiome_panel, fig.height = 22, fig.width = 17}
## Arrange the figures
arrange_row1 <- ggarrange(adiv_outcome_plot, 
                          bdiv_outcome_plot,
                          ncol = 2, 
                          labels = paste0("(", letters[1:2], ")"), 
                          font.label = list(size = 14, 
                                            face = "bold", 
                                            family = "serif"))


arrange_row2 <- ggarrange( 
                          phylum_abund_outcome_p, 
                          genus_abund_outcome_p,
                          ncol = 2, 
                          labels = paste0("(", letters[3:4], ")"),
                          font.label = list(size = 14, 
                                            face = "bold", 
                                            family = "serif"),
                          common.legend = TRUE,
                          legend = "bottom")

## edit bottom margin to allow more space for common legend
arrange_row2 <- arrange_row2 +
  theme(plot.margin = margin(
    t = 0.2,
    r = 0.2,
    b = 0.4,
    l = 0.2,
    unit = "inches"
  ))

## Create an empty plot to use as a spacer
empty_plot <- ggplot() + theme_void()

arrange_row3 <- ggarrange(empty_plot, 
                          fbratio_outcome_plot,
                          empty_plot,
                          venn_p,
                          ncol = 4, 
                          labels = c("", "(e)", "", "(f)"), 
                          font.label = list(size = 14, 
                                            face = "bold", 
                                            family = "serif"),
                          widths = c(0.2, 0.5, 0.2, 1))

microbiome_panel <- ggarrange(arrange_row1,
                              arrange_row2,
                              arrange_row3,
                              nrow = 3,
                              heights = c(0.8,1,0.9))

## view 
microbiome_panel

## Save
ggsave(
  filename = paste0(
    analysis_dir,
    "figs/HIC_microbiome_heat_tolerance_panel",
    ".png"
  ),
  plot = microbiome_panel,
  height = 22,
  width = 17,
  unit = "in",
  dpi = 600
)

## Save
ggsave(
  filename = paste0(
    analysis_dir,
    "figs/HIC_microbiome_heat_tolerance_panel",
    ".tiff"
  ),
  plot = microbiome_panel,
  height = 22,
  width = 17,
  unit = "in",
  dpi = 600
)

## Save
ggsave(
  filename = paste0(
    analysis_dir,
    "figs/HIC_microbiome_heat_tolerance_panel",
    ".pdf"
  ),
  plot = microbiome_panel,
  height = 22,
  width = 17,
  unit = "in",
  dpi = 600
)
```

## Control Panel

```{r Section7_pos_neg_control_panel, fig.height = 10, fig.width = 18.5}
## create positive and negative control panel
P_N_con_panel <- ggarrange(PC_plot, 
                           PC_plot_theo,
                           NC_plot_caption,
                           labels = LETTERS[1:3],
                           font.label = list(size = 18, 
                                             face = "bold", 
                                             family = "serif"),
                           vjust = 1.0,
                           ncol = 3, 
                           nrow = 1)

## view 
P_N_con_panel
```


# Session Info
``` {r sessionInfo}
sinfo <- sessionInfo()
sinfo
save.image(paste0(analysis_dir, "/HIC_16S_analysis_manuscript_heat_tolerance.Rout"))
```
